### 2017-04-04 proposal/proposal.tex
proposal document

\documentclass[11pt,twoside]{article}
\usepackage{etex}
\newcommand{\num}{6{} }

\raggedbottom

%geometry (sets margin) and other useful packages
\usepackage{geometry}
\geometry{top=1in, left=1in,right=1in,bottom=1in}
 \usepackage{graphicx,booktabs,calc}

%=== GRAPHICS PATH ===========
%\graphicspath{{./140408-Images/}}
% Marginpar width
%Marginpar width
\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } } 
\setlength{\marginparwidth}{.5in}
%\reversemarginpar
%\setlength{\marginparsep}{.02in}

%% Fonts
% \usepackage{fourier}
% \usepackage[T1]{pbsi}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{minted}

\usepackage{rotating}
%% Cite Title
\usepackage[style=authoryear,natbib,maxcitenames=2,doi=false,isbn=false,url=false,eprint=false]{biblatex}
\addbibresource{bib/references.bib}

%%% Counters
\usepackage{chngcntr,mathtools}
\counterwithout{figure}{section}
\counterwithout{table}{section}

\numberwithin{equation}{section}

%% Captions
\usepackage{caption}
\captionsetup{
  labelsep=quad,
  justification=raggedright,
  labelfont=sc
}

%AMS-TeX packages
\usepackage{amssymb,amsmath,amsthm} 
\usepackage{bm}
\usepackage[mathscr]{eucal}
\usepackage{colortbl}
\usepackage{color}

\usepackage{epstopdf,subfigure,hyperref,enumerate,polynom,polynomial}
\usepackage{multirow,minitoc,fancybox,array,multicol}

\definecolor{slblue}{rgb}{0,.3,.62}%
%\renewcommand{\chaptermark}[1]{ \markboth{#1}{} }
\renewcommand{\sectionmark}[1]{ \markright{#1}{} }

\usepackage{fancyhdr}
\pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
%\fancyheadoffset{30pt}
%\fancyfootoffset{30pt}
\fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
\fancyhead[RO,LE]{\small Page \thepage} 
\fancyfoot[RO,LE]{\small }% PR \num S-2015} 
\fancyfoot[LO,RE]{\small }%\scshape MODL} 
\cfoot{} 
\renewcommand{\headrulewidth}{0.1pt} 
\renewcommand{\footrulewidth}{0pt}
%\setlength\voffset{-0.25in}
%\setlength\textheight{648pt}

\usepackage{paralist}

%%% FORMAT PYTHON CODE
\usepackage{listings}
% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8}  % for normal

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}

%\usepackage{listings}

% % Python style for highlighting
% \newcommand\pythonstyle{\lstset{
% language=Python,
% basicstyle=\footnotesize\ttm,
% otherkeywords={self},             % Add keywords here
% keywordstyle=\footnotesize\ttb\color{deepblue},
% emph={MyClass,__init__},          % Custom highlighting
% emphstyle=\footnotesize\ttb\color{deepred},    % Custom highlighting style
% stringstyle=\color{deepgreen},
% frame=tb,                         % Any extra options here
% showstringspaces=false            % 
% }}

% % Python environment
% \lstnewenvironment{python}[1][]
% {
% \pythonstyle
% \lstset{#1}
% }
% {}

% % Python for external files
% \newcommand\pythonexternal[2][]{{
% \pythonstyle
% \lstinputlisting[#1]{#2}}}

% % Python for inline
% \newcommand\pythoninline[1]{{\pythonstyle\lstinline!#1!}}

\newcommand{\osn}{\oldstylenums}
\newcommand{\dg}{^{\circ}}
\newcommand{\lt}{\left}
\newcommand{\rt}{\right}
\newcommand{\pt}{\phantom}
\newcommand{\tf}{\therefore}
\newcommand{\?}{\stackrel{?}{=}}
\newcommand{\fr}{\frac}
\newcommand{\dfr}{\dfrac}
\newcommand{\ul}{\underline}
\newcommand{\tn}{\tabularnewline}
\newcommand{\nl}{\newline}
\newcommand\relph[1]{\mathrel{\phantom{#1}}}
\newcommand{\cm}{\checkmark}
\newcommand{\ol}{\overline}
\newcommand{\rd}{\color{red}}
\newcommand{\bl}{\color{blue}}
\newcommand{\pl}{\color{purple}}
\newcommand{\og}{\color{orange!90!black}}
\newcommand{\gr}{\color{green!40!black}}
\newcommand{\nin}{\noindent}
\newcommand{\la}{\lambda}
\renewcommand{\th}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\G}{\Gamma}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,thick,inner sep=1pt] (char) {\small #1};}}

\newcommand{\bc}{\begin{compactenum}[\quad--]}
\newcommand{\ec}{\end{compactenum}}

\newcommand{\p}{\partial}
\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\dpd}[2]{\dfrac{\partial{#1}}{\partial{#2}}}
\newcommand{\pdd}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Virtual city generation}
%\author{Jimi Oke}
\date{\small\today}
\maketitle

\thispagestyle{empty}

\tableofcontents
\section{Objective}

\section{Network synthesis approach}
We propose using the Latent Space modeling approach for generating our representative street networks.
The Latent Space Model was introduced by \textcite{hoff2002latent} for modeling social networks.
The formulation was further extended by \cite{zhou2015generating} to develop urban network models.
Essentially, given a relationship specifications between actors $y_{i,j}$ in a network at positions $z_i$ and $z_j$,
then the probability of pairwise connections is
\begin{equation}
  \label{eq:1}
  P(Y, Z, \theta) = \prod_{i\ne j} P(y_{i,}|z_i, z_j, x_{i,j},\theta)
\end{equation}
where $x_{i,j}$ are possible covariates for pairing tendencies.
This probability is parametrized by a logit model, which, for now, ignoring any covariates, is given by
\begin{equation}
  \label{eq:2}
  P(y_{i,j}|z_i,z_j,\theta) = \frac{1}{1+e^{\lambda D_{i,j}}}
\end{equation}
where $D_{i,j}$ is the distance metric specified.

For urban street networks, the Euclidean distance (i.e.\ $|z_i - z_j|$) is not the only measure of interest.
The accessibility of one node from the other is of overall importance and one means of capturing this is by using the shortest distance between the pair of nodes in the network.

Further, the distance or similarity matrix must be generated on a latent space using a kernel approach in order to reduce dimensionality.
A readily applicable one is multi-dimensional scaling (MDS).
Similar techniques, such as exploratory factor analysis or auto-encoding approaches may be applied to uncover the latent structure in the pair-wise nodal relationships.
The similarity matrix in the latent space can thus be denoted $D_{i,j}^L$.

Following the approach of \cite{zhou2015generating}, the model can be evaluated and tuned (via the parameter $\lambda$) by comparing the following properties of the synthesized networks to the actual ones in our clusters:
\begin{itemize}
\item connectivity
\item diameter
\item path length
\item triads, clustering coefficient
\item degree distributions
\end{itemize}

Besides accessibility, we will also explore the encoding of other information into the latent space model.

% \section{}
% \subsection{Pump price of gasoline}
% Currently, it appears there is no readily available source of data on gas prices
% by city on a global scale. However, there are multiple sources on current
% gasoline prices on a country basis from
% \href{http://www.globalpetrolprices.com/gasoline_prices/}{Global Petrol
%   Prices}. Prices for cities in these countries Further disaggregation for
% US/Canadian cities is available from
% \href{http://then.gasbuddy.com/GB_Price_List.aspx}{GasBuddy.com}.

%\subsection{Modeshare}

\printbibliography

\end{document}

### 2017-04-04 roadNets.py

@@ -57,7 +57,7 @@ def getGraph(location, type='drive'):
 	# gdf = ox.gdf_from_place(location)
 	# areaSqM = ox.project_gdf(gdf).unary_union.area
 	try:
-        G = ox.graph_from_place(location, network_type='drive')
+		G = ox.graph_from_place(location, network_type='drive')
 		ox.plot_graph(ox.project_graph(G))
 	except ValueError:
 		G = ox.graph_from_place(location, network_type='drive',which_result=2)
@@ -97,7 +97,6 @@ def writeToDataFrame(stats, index):
 def main():
 
 	g = getGraph('Mountainside, NJ, USA', type='drive')
-	return g
 
 	# G = ox.graph_from_place('Manhattan, New York, USA', network_type='drive')
 	# ox.plot_graph(ox.project_graph(G))

### 2017-08-15 review/review.tex
renamed file

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{listings}
\usepackage[english]{babel}
\usepackage{enumerate}
\lstset{frame=tlrb,xleftmargin=\fboxsep,xrightmargin=-\fboxsep}

\usepackage[style=numeric, backend=biber, maxcitenames=2]{biblatex}
\bibliography{bib/references}

\begin{document}

\title{Population Synthesis Literature Review}
\author{Jin Zhou}
\maketitle

\section{Goal}
Since the step of urban typology has been done, we need to fill in the blank between city clustering results and a complete SimMobility\cite{adnan2016simmobility} input of virtual city.
The city model that SimMobility takes as input has two components.
One is the demand and the other is the supply, which includes the road and transit network.
To model the demand, a methodology has to be developed to construct the population with desired demographics and allocate the population in space.
Next to it, we also need to figure out how to use the synthesized households and individuals in disaggregate level to get the travel demand in the city.
The literature review at this time is to gain a better idea of previous efforts to address the problem of population synthesis and assignment.

\subsection{Questions:}
\begin{enumerate}
	\item What should the input of travel demand generation look like? This guides the adaptation of population synthesis method from literature.
	\item What attributes should be included to characterize the population? Age, sex, educational level, and there must be some attributes reflect the traveling behavior of people. 
	\item Do we synthesize households or personnel? Or how to match individuals to households?
	\item We need to come up with a solution to cities don't have PUMS data and virtual cities that don't even have census data for demographics.	
\end{enumerate}

\section{Prior Work}

Household assignment and spatial allocation are two big issues we need to solve when facing the problem of population synthesis.
Specifically, the first one addresses the problem of how to match individuals to the household where they belong, while the spatial allocation distributes households in geographical space.
It is necessary because the ultimate goal of our work is to do transportation simulation, which can only be achieved by knowing the location of each agent. \\

There has been lots of literature on population synthesis.
The methods used varies from Iterative Proportional Fitting (IPF), MCMC-based sampling \cite{farooq2013simulation}, to Bayesian network approach \cite{sun2015bayesian}. The original work in this field is presented in \cite{beckman1996creating}. It used a method called Iterative Proportional Fitting (IPF) to synthesize population based on data from two sources. What they do is iteratively fitting the cells of a multi-way contingency table expanded by attributes that characterize a household/individual so that it satisfies the constraints of marginal distribution of demographics and also preserves the correlation structure carried by sample data of the whole population. Here the marginals of demographics comes from census data in the level of census tract or block group while the correlation structure need to be derived from survey data of target population. In case of the absence of disaggregated sample data, IPF can still give a solution by iteratively adjust the cells in accordance only with the marginal constraints. However, lots of different joint distributions can share the same marginals and IPF may converge to anyone of them. This is certainly a downside of traditional IPF, and is also the reason why we need the correlation structure of sample data to make the solution captures the true population distribution. After the fitting of the table, a desired number of synthetic population is sampled according to individual or household weights in the table. And this is where the sensitivity of IPF to the quality of data and the sample size stems from. In this method, spatial allocation is solved concurrently with the population synthesis, since it generates the population for each zone and then adds up to the whole population. On contrary to this, one can also synthesize the total population first and then allocate them to different zones. However, this includes implementing a detailed matching method between households and zones, like what's developed in \cite{ge2014virtual}, which is also very time-consuming.

Ever since Beckham et al. used IPF to do population synthesis, limitation of this method and the targeted modification has been published by other researchers. By IPF alone, one can only separately generate a set of weights either for individual or household. According to \cite{ye2009methodology}, the synthetic population generated based on the application of household weights can give a joint distribution of person attributes far from the given person marginal distribution.
This is due to the fact that all persons are just simply selected from chosen households according to household weights.
\textcite{ye2009methodology} has improved this problem of IPF by a method called Iterative Proportional Updating (IPU), and they claim that the algorithm they proposed can match the person-level distribution and household-level attributes as closely as possible. \\

MCMC-based sampling is brought up by \textcite{farooq2013simulation}.
The problem is defined as how to use partial views, including samples and conditional distributions, to estimate the true joint distribution of population attributes, where the attributes are defined as $X = (X^1, X^2, ... X^n)$.
Here, they used Gibbs Sampling to generate certain amount of population based on prior knowledge of the joint distribution and proved that the sampled population is very much similar to the true simulation in the statistical sense.
It has been pointed out that, since the task of getting full conditionals won't be trivial in reality, they replaced some of conditional distributions by marginal distributions.
Thus, if one wants to implement MCMC, the sample data of the entire population is fundamental.
Furthermore, the derivation of conditional distribution is not trivial especially when the number of attributes gets large.

The population synthesis method that \textcite{petrik2016measuring} used adapted ideas from probabilistic graphical models (PGM).
In their graphical model, there are 7 socio-demographic attributes, which are Age, Gender, Employment, Income, Car ownership, Purpose of trip, and Who pay for the trip, to represent a person.
Once the dependencies between variables are specified, we need to calculate or estimate conditional probability of one node given its parents.
The flexibility of this method is that those conditional probabilities can be derived from different sources like census data or surveys on individuals.
Finally, the synthetic population is just random draws from the probabilistic graphical model.
Though we can take marginal distributions of all variables as supplement for the graphical model, the population generated by Bayesian network model can only precisely match the marginals of mother nodes.
This is because the distribution of a node is uniquely defined by its parent node and the conditional distribution based on PGM.
And this point should be taken into consideration when choosing the population synthesis method.

\section{Data}
Most paper in population synthesis used the Public Use Microdata Sample (PUMS), which is available both in individual level and household level.
It includes multiple socio-economic parameters of a household or an individual in the region.
The sample rate of this dataset is about 1\% to 10\% of the total population in the region.
We found PUMS data is complete in US cities but haven't found microdata in other cities in the world.
Currently, we have found aggregated information of population, age, sex, household, education in districts in Beijing.

The choice of algorithm should be based both on the data we have and the appearance of the final result we want.
What is also important for us is the simplicity and novelty of methodology.

\section{Iterative Proportional Fitting}

Using IPF to synthesize population includes two major steps.
In the first step, the disaggregate samples of population are used to initialize the contingency table expanded by attributes, and the fitting is carried out to meet marginal constraints from aggregated data.
In the next step, the fitted contingency table is used to sample the population.
To describe IPF, we need to define the contingency table it fits.
Suppose a person or household is described by \textit{m} attributes (or demographics), then we need to develop a m-way table.
For each demographic $i$, there are $n_i$ categories.
Thus, a cell $(i_1, i_2,..., i_m)$ represents a kind of person/household, where $i_j = 1,2,...n_j$ is the observed value of the $j$-th demographic with $n_j$ categories it can choose from.
From sample data like PUMS, $n$ is the total number of observations while $n_{i_1, i_2,...,i_m}$ is the counts for people/household of this type.
Hence, the proportion of this type in the sample is denoted by
\begin{equation}
  \label{eqn:prop}
   p_{i_1,i_2,...,i_m} = \frac{n_{i_1, i_2,...,i_m}}{n}
\end{equation}
The constraints is defined as $T_k^j$, which is the marginal totals for the $k$-th category of attribute $i$ from census data.
We can easily get the equation
\begin{equation}
  \label{eq:2}
\sum_i\sum_{k=1}^{n_i}T_k^j=n.  
\end{equation}
Let $p_{i_1,i_2,...,i_m}^{(t)}$ denote estimated proportion of cell $(i_1, i_2,..., i_m)$ in iteration $t$ and
\begin{equation}
  \label{eq:3}
  p_{...,i_j=k,...} = \sum_{i_1=1}^{n_1}...\sum_{i_m=1}^{n_m}p_{i_1,i_2,...,i_j=k,...,i_m}^{(t)}  
\end{equation}

The IPF starts with initiating all weights in the contingency table by  
\begin{equation}
	p_{i_1,i_2,...,i_m}^{(0)} = p_{i_1,i_2,...,i_m}
\end{equation}
and in each iteration, the proportions are updated by the formula
\begin{equation}
  p_{i_1,i_2,...,i_j=k,...,i_m}^{(new)} = p_{i_1,i_2,...,i_j=k,...,i_m}^{(old)} * \frac{T_k^j * p_{...,i_j=k,...}^{(new)}}{n} 
\end{equation} for all categories of each attribute, until it converges.
Within an iteration, $p^{(old)}$ for the first marginal corresponds to $p^{(t-1)}$, which is the result from last iteration.
And for marginals come next, $p^{(old)}$ should be set to proportions calculated from previous marginals.
According to \cite{beckman1996creating}, the algorithm converges in 10 to 20 iterations.

In the next step, the synthetic population shall be constructed by selecting households from PUMS in proportion to the estimated weights in the multiway table.
Specifically, the number of households of each demographic type in a census tract can be obtained by multiplying the total number of households by the probability in the cell representing this household type, or by drawing the numbers at random according to these probabilities.
However, the household type represent by a cell in multiway table is not exactly that of PUMS data.
The fitted multiway table can only be span by control variables, which is defined as variables which have marginal distribution in census data.
However, households and individuals in disaggregate data are usually described by additional desired variables.
So, the probability of a household type in PUMS data being chosen shall be assigned according to the distance between such a household $p$ and a household type $c$ in multi-way table.
In \cite{beckman1996creating}, this probability is defined as 
\begin{equation}
	D(p,c) = w_p \prod_{i \in J}\Big(1-|(d_i^p-d_i^c)/r_i|^k\Big)*\prod_{i \notin J}\Big(1-\Delta(d_i^p, d_i^c)\Big)
\end{equation}

\section{Proposed method}

\begin{enumerate}[Step 1:]
\item Network: modify OSM network file, make it legal input to SimMobility
  \begin{enumerate}[1.]
  \item links has to be unidirectional
  \item shorten the links and add turning points
  \item maintain the attributes of each object in the network
  \end{enumerate}
\item Population Synthesis
  \begin{enumerate}[1. ]
  \item Data collection: survey data, aggregate census data for zones in each city
  \item Learn the parameter and the structure of Bayesian Network which captures the structure of population
  \item sampling from Bayesian Network to get whole population in the city
  \item Clustering the synthetic population in a city by K-means
  \item Divide the city into zones according to clustering result, and form the aggregate census data for zones
  \end{enumerate} 
\end{enumerate}

Although applying IPF to population synthesis is prevailing and the idea is straightforward, the algorithm itself is time consuming and expansive once the combinations of demographics gets large.
Suppose we have 2 gender categories, 7 age groups, 5 income groups, 5 levels of education, 9 types of occupation, 2 kinds of car ownership, and 5 household types.
These alone give rise to $31\,500$ types of individuals, not to mention describing individuals with more demographics.
Another issue with published methods is that the time we spend to generate population for one zone has to be multiplied by the number of zones in a city in order to get the whole population for the city, which means the time for a complete process is hundreds of that for a single zone if the zone is in the sense of census tract.
Since what we want is something less time consuming, a new methodology has to be developed.

Considering the accessibility of the two sources of data that the synthesizer needs, one can always get sampled survey data in a city level but not necessarily census data for zones in a city.
Given the global scope of our problem, there are cities whose zones are even not defined properly.
Thus, we will face the problem of defining the zones of a city by our own in case of the data is in shortage.
In developing the zones in a city, we will borrow the idea of zones from \textcite{fielbaum2016parametric}, where a city is divided into one CBD and subcenters along with their peripheral.
It has to be admitted that in this way, we will not have aggregate census data for zones. This is a problem we will address later.

\subsection{Two approaches to spatial assignment of households}
\subsubsection{K-means clustering}
With regard to the assignment of populations to zones, we will cluster the population at city level with K-means first.
We reason that similar and highly correlated people will most likely live in the same area within a city.
Then we obtain grouped people with similar attributes.
We choose K-means because it is more intuitive if we want a spatial allocation and the feature of hidden centroid for each cluster will hopefully capture the characteristic of the zone.

\subsubsection{Zonal disaggregation and random sampling}
\textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
First, they obtain land-use data, which they use to disaggregate zonal characteristics (e.g.\ household location or firm location) into raster cells.
The size of these cells are chosen based on computational tractability and desired resolution.
Weights are then assigned to the cells based on the corresponding land-use category.
For the household example, \textcite{moekel2003microsimulation} aggregate land-use attributes into the following: high-density housing, low-density housing, industry and open space;
to which weights of 10, 5, 1 and 1 are assigned, respectively.
The weights are then normalized by zone, and households are randomly selected to populate each cell to fit each proportion.
The address of each household is assumed to be the centroid of each cell, which each associated with its parent zone.
The rasterization is performed by a so-called ``point-in-polygon'' algorithm.
This approach presents a fast way for household assignment, assuming prior zonal assignment is available.
The populations themselves can first be generated using a simple IPF/U routine.

\newpage
\section{References}
\printbibliography[heading=none]
\end{document}

### 2017-08-17 population/houston_puma.py
archive previous work

@@ -1,0 +1,11 @@
+
+
+a = ['04503', '04504']
+for i in range(1, 38):
+	if i < 10:
+		a.append('0460'+str(i))
+	else:
+		a.append('046'+str(i))
+a.extend(['04702','04801','04901','04902','04903','04904','04905'])
+print a
+print len(a)

### 2017-08-17 population/microdata_eda.R

@@ -1,186 +1,0 @@
-
-
-library(ipfp)
-library(plyr)
-library(bnlearn)
-library(ggplot2)
-library(data.table)
-# individual variables
-
-# household number: pcode
-# people in HH: p102
-# relationship to the head of household: p103 
-# Delete observations without "Urban Hukou": p104
-# Sex: p105
-# Age, need to aggregate: p106
-# Status (employment): p107
-# marriage: p109
-# education level: p112
-# total income: p201
-
-
-
-# household variables
-# total financial assets: h401
-# location of residence: h501
-# current household living standard group in city: h604a
-
-
-individual <- read.table('data/21741-0001-Data.tsv', header=T, sep='\t')
-household <- read.table('data/21741-0002-Data.tsv', header=T, sep='\t')
-
-
-# individual <- individual[individual$city == '110101']
-# select people in city
-
-households_tb_kept <- individual[individual$p104 == 1, 'pcode']
-household <- subset(household, pcode %in%  households_tb_kept)
-individual <- individual[individual$p104==1 & !is.na(individual$pcode),]
-
-
-dim(individual)
-colnames(individual)
-count(individual, 'city')
-plot(count(individual,'p106'),cex=0.5)
-
-
-inRawData <- as.data.frame(individual[,c('pcode','p102','p103','p105','p106','p107','p109','p112','p201')])
-#inRawData <- individual[,c('pcode','p105','p106','p107','p109','p112','p201')]
-hhRawData <- as.data.frame(household[,c('pcode','h401','h501')])
-
-
-colnames(inRawData) <- c('pcode','hh_size','rel','sex','age','emp_status','marriage','edu','inc')
-colnames(hhRawData) <- c('pcode','fin_asset','loc_res')
-
-# hh_size:
-# categories: 1,2,3,4, >5
-for (i in unique(inRawData$pcode)){
-  inRawData[inRawData$pcode == i, 'hh_size'] <- max(inRawData[inRawData$pcode == i,'hh_size'])
-}
-inRawData$hh_size <- ifelse(inRawData$hh_size >= 5, 5, inRawData$hh_size)
-
[…]

### 2017-08-17 population/houston_puma.py
pums preprocessing part 1

@@ -1,11 +1,61 @@
 
+import pandas as pd 
+import os
+import ipfn
 
-a = ['04503', '04504']
-for i in range(1, 38):
-	if i < 10:
-		a.append('0460'+str(i))
-	else:
-		a.append('046'+str(i))
-a.extend(['04702','04801','04901','04902','04903','04904','04905'])
-print a
-print len(a)
+
+def preprocessingPersonPUMS(dirname,filename,pumaInArea):
+	'''
+	delete unnecessary columns
+	and aggregate variables like age into categories
+
+	df: Data Frame - original PUMS data
+	pumaInArea: list - all related puma
+	'''
+
+	df = pd.read_csv(os.path.join(dirname,filename))
+	print df.shape
+	# Specify all variable names that is related
+	df = df[df['PUMA'].isin(map(int,pumaInArea))][['SERIALNO','SPORDER','PUMA','AGEP',
+	'COW','JWMNP','JWTR','MAR','RELP','SEX','SCHL','ESR','PINCP','ADJINC']]
+	print df.shape
+	df.to_csv('data/ss15ptx_clean.csv', index = False)
+
+	return df
+
+
+
+
+
+def preprocessingMarginal():
+	pass
+
+dirname = 'data'
+hhfilename = 'ss15htx.csv'
+psfilename = 'ss15ptx.csv'
+map_filename = '2010_Census_Tract_to_2010_PUMA.txt'
+
+
+# Give all the counties in A specific metro area
+# Here we are in Texas (48), Houston metro area
+stateID = '48'
+countyInArea = ['015','039','071','157','167','201','291','339','473']
+# Read the map file
+mapCTtoPUMA = pd.read_csv(os.path.join(dirname,map_filename),sep = ',',converters={'STATEFP':str,'COUNTYFP':str,'PUMA5CE':str})
+
+
+# Get PUMA which is in Houston metro area
+pumaInArea = mapCTtoPUMA[mapCTtoPUMA['STATEFP']==stateID]
+pumaInArea = pumaInArea[pumaInArea['COUNTYFP'].isin(countyInArea)]['PUMA5CE'].unique()
+print pumaInArea
+
+
+
+
+#household = pd.read_csv(os.path.join(dirname,hhfilename))
+# Write a subset of the whole data, easier to open in excel
+#household[0:500].to_csv(os.path.join(dirname,'sample_'+hhfilename),index = False)
+
[…]

### 2017-08-17 population/marginal_constraints.R

@@ -1,25 +1,0 @@
-
-a <- t(read.table('sex_age'))
-a <- cbind(a, rep(0:22,each=3))
-a <- cbind(a, rep(0:2,23))
-
-a <-data.frame(a)
-colnames(a) <- c('pop','age_category','sex')
-b <- a[a$age_category!=0 & a$sex!=0,]
-
-b[3,'pop'] <- b[3,'pop']+b[1,'pop']
-b[4,'pop'] <- b[4,'pop']+b[2,'pop']
-b <- b[-c(1,2),]
-
-sex_age <- array(NA, c(7,2))
-for (i in 1:6){
-  sex_age[i,1] <-b[4*i-3,1]+b[4*i-1,1]
-  sex_age[i,2] <-b[4*i-2,1]+b[4*i,1] 
-}
-
-sex_age[7,1] <- sum(b[c((12:20)*2+1),1])
-sex_age[7,2] <- sum(b[c((13:21)*2),1])
-
-
-sd <- t(read.table('sex_edu'))
-sd <- t()

### 2017-08-17 population/houston_puma.py
preprocessing household data

@@ -15,7 +15,7 @@ def preprocessingPersonPUMS(dirname,filename,pumaInArea):
 
 	df = pd.read_csv(os.path.join(dirname,filename))
 	print df.shape
-	# Specify all variable names that is related
+	# Specify all person variable names that is related
 	df = df[df['PUMA'].isin(map(int,pumaInArea))][['SERIALNO','SPORDER','PUMA','AGEP',
 	'COW','JWMNP','JWTR','MAR','RELP','SEX','SCHL','ESR','PINCP','ADJINC']]
 	print df.shape
@@ -24,8 +24,17 @@ def preprocessingPersonPUMS(dirname,filename,pumaInArea):
 	return df
 
 
+def preprocessingHouseholdPUMS(dirname, filename, pumaInArea):
+	
+	df = pd.read_csv(os.path.join(dirname, filename))
+	print df.shape
+	# Specify all household variable names that is related
+	df = df[df['PUMA'].isin(map(int,pumaInArea))][['SERIALNO','PUMA','NP','VEH','HHT',
+	'HINCP','HUPAC','WIF']]
+	print df.shape
+	df.to_csv('data/ss15htx_clean.csv', index = False)
 
-
+	return df
 
 def preprocessingMarginal():
 	pass
@@ -52,9 +61,10 @@ print pumaInArea
 
 
 
-#household = pd.read_csv(os.path.join(dirname,hhfilename))
+household = preprocessingHouseholdPUMS(dirname, hhfilename, pumaInArea)
 # Write a subset of the whole data, easier to open in excel
-#household[0:500].to_csv(os.path.join(dirname,'sample_'+hhfilename),index = False)
+# household[0:500].to_csv(os.path.join(dirname, 'sample_'+hhfilename), index = False)
+# household[housthold['PUMA'].isin(pumaInArea)].to_csv(os.path.join(dirname,hhfilename+'_clean'),index = False)
 
 person = preprocessingPersonPUMS(dirname, psfilename, pumaInArea)
 #person[0:500].to_csv(os.path.join(dirname, 'sample_'+psfilename), index = False)

### 2017-10-17 pop-synth/beijing_bn/microdata_eda.R
edits to documentation

@@ -1,0 +1,191 @@
+
+
+library(ipfp)
+library(plyr)
+library(bnlearn)
+library(ggplot2)
+library(data.table)
+# individual variables
+
+# household number: pcode
+# people in HH: p102
+# relationship to the head of household: p103 
+# Delete observations without "Urban Hukou": p104
+# Sex: p105
+# Age, need to aggregate: p106
+# Status (employment): p107
+# marriage: p109
+# education level: p112
+# total income: p201
+
+
+
+# household variables
+# total financial assets: h401
+# location of residence: h501
+# current household living standard group in city: h604a
+
+
+individual <- read.table('data/21741-0001-Data.tsv', header=T, sep='\t')
+household <- read.table('data/21741-0002-Data.tsv', header=T, sep='\t')
+
+
+# individual <- individual[individual$city == '110101']
+# select people in city
+
+households_tb_kept <- individual[individual$p104 == 1, 'pcode']
+household <- subset(household, pcode %in%  households_tb_kept)
+individual <- individual[individual$p104==1 & !is.na(individual$pcode),]
+
+
+dim(individual)
+colnames(individual)
+count(individual, 'city')
+plot(count(individual,'p106'),cex=0.5)
+
+
+inRawData <- as.data.frame(individual[,c('pcode','p102','p103','p105','p106','p107','p109','p112','p201')])
+#inRawData <- individual[,c('pcode','p105','p106','p107','p109','p112','p201')]
+hhRawData <- as.data.frame(household[,c('pcode','h401','h501')])
+
+
+colnames(inRawData) <- c('pcode','hh_size','rel','sex','age','emp_status','marriage','edu','inc')
+colnames(hhRawData) <- c('pcode','fin_asset','loc_res')
+
+# hh_size:
+# categories: 1,2,3,4, >5
+for (i in unique(inRawData$pcode)){
+  inRawData[inRawData$pcode == i, 'hh_size'] <- max(inRawData[inRawData$pcode == i,'hh_size'])
+}
+inRawData$hh_size <- ifelse(inRawData$hh_size >= 5, 5, inRawData$hh_size)
+
[…]

### 2017-10-17 pop-synth/categorize.py

@@ -1,0 +1,214 @@
+
+import pandas as pd 
+import numpy as np 
+import os
+import math
+
+def categorizePersonPUMS(dirname, psfilename):
+	'''
+	replace values of each variable in sample data with categories from aggregate data
+
+	'''
+
+	df = pd.read_csv(os.path.join(dirname, psfilename))
+	selectColumns = ['SERIALNO','SPORDER','PUMA','SEX','AGEP','MAR','SCHL','PINCP']
+	df = df[selectColumns]
+	df.columns.values[3:] = ['sex','age','mar','edu','inc']
+	print max(df['age'])
+	print min(df['inc']), max(df['inc'])
+
+	sexC = {
+			'Male': [1],
+			'Female': [2]
+			}
+
+	ageC = {
+			'Under 5 years': range(0,5),
+			'5 to 17 years': range(5,18),
+			'18 to 24 years': range(18,25),
+			'25 to 44 years': range(25,45),
+			'45 to 54 years': range(45,55),
+			'55 to 64 years': range(55,65),
+			'65 to 74 years': range(65,75),
+			'75 years and over': range(75,125)
+			}
+	marC = {
+			'Never married': [5],
+			'Married': [1],
+			'Divorced or separated': [3,4],
+			'Widowed': [2]
+			}
+
+	eduC = {
+			'Less than high school graduate': range(1,16),
+			'High school graduate': [16,17],
+			'Some college or associate\'s degree': range(18,21),
+			'Bachelor\'s degree': [21],
+			'Graduate or professional degree': range(22,25),
+			}
+
+	incC = {
+			'$1 to $9,999 or loss': [min(min(df['inc']),1), 9999],
+			'$10,000 to $14,999': [10000,14999],
+			'$15,000 to $24,999': [15000,24999],
+			'$25,000 to $34,999': [25000,34999],
+			'$35,000 to $49,999': [35000,49999],
+			'$50,000 to $64,999': [50000,64999],
+			'$65,000 to $74,999': [65000,74999],
+			'$75,000 or more': [75000, max(max(df['inc']),80000)],
+			}
+
+
+
+	categories = {'sex': sexC, 'age': ageC, 'mar': marC, 'edu': eduC, 'inc': incC}
+	newdf = df.copy()
+	# processing sex, age, marriage status
+	for sub in ['sex','age','mar']:
+		for cat in categories[sub].keys():
+			newdf[sub] = newdf[sub].replace(categories[sub][cat],cat)
+
[…]

### 2017-10-17 pop-synth/draw.py

@@ -1,0 +1,125 @@
+import pandas as pd
+import numpy as np
+from collections import OrderedDict
+import itertools
+import os
+
+
+def simple_draw(tot,weights):
+
+
+	'''
+	Input:
+	weights - is MultiIndex Series with index indicates the categories of variables,
+			  the data gives the weight of this combination of categories
+
+	'''
+
+	p = weights/weights.sum()
+	index = weights.index.values.tolist()
+	print index
+	return np.random.choice(len(index), size = tot, p = p, replace = True)
+
+
+
+def draw_synth_hh(ipf_results, hh_sample, ps_sample,mapCTtoPUMA, countyTable, out_dir):
+	'''
+	Input:
+	ipf_results: dictionary with keys represent a county
+	ipf_results[counyt]: dataframe, first columns record categories for variables 
+						 while the last column 'total' is the weight for this combination
+	hh_sample: dataframe, first columns are 'SERIALNO','PUMA'
+	ps_sample: dataframe, first columns are 'SERIALNO','SPORDER','PUMA'
+	
+	do MC sampling based on weights in ipf result
+
+	'''
+
+	synth_hh = []
+	synth_ps = []
+	i = 0
+	for county in ipf_results.keys():
+		i += 1
+		# modify index for weights
+		ipf_result = ipf_results[county]
+		index_title = list(ipf_result.columns.values[:-1])
+		index_title.sort()
+		tot = sum(ipf_result['total'])
+
+		# select sample in this county
+		countyId = countyTable.loc[countyTable['name'] == county,'id'].values[0]
+		pumaInCounty = mapCTtoPUMA.loc[mapCTtoPUMA['COUNTYFP']==countyId,'PUMA5CE']
+		pumaInCounty = [int(x) for x in pumaInCounty.tolist()]
+		# household and person sample in the county
+		hhspc = hh_sample[hh_sample['PUMA'].isin(pumaInCounty)]
+		psspc = ps_sample[ps_sample['PUMA'].isin(pumaInCounty)]
+
+		# merge weights of each type of household with household sample data
+		merge_df = pd.merge(hhspc,ipf_result, on=index_title)
[…]

### 2017-10-17 pop-synth/houston_puma.py

@@ -1,0 +1,264 @@
+
+import pandas as pd 
+import numpy as np
+import os
+from categorize import *
+from marginal_processing import *
+from ipf import *
+from draw import *
+
+
+
+def selectPersonPUMS(dirname,filename,pumaInArea):
+	'''
+	delete unnecessary columns
+
+	df: Data Frame - original PUMS data
+	pumaInArea: list - all related puma
+	'''
+
+	df = pd.read_csv(os.path.join(dirname,filename))
+	print df.shape
+	# Specify all person variable names that is related
+	df = df[df['PUMA'].isin(map(int,pumaInArea))][['SERIALNO','SPORDER','PUMA','AGEP',
+	'COW','JWMNP','JWTR','MAR','RELP','SEX','SCHL','ESR','PINCP','ADJINC']]
+	print df.shape
+	df.to_csv('data/ss15ptx_clean.csv', index = False)
+
+
+
+def preprocessingHouseholdPUMS(dirname, filename, pumaInArea):
+	
+	df = pd.read_csv(os.path.join(dirname, filename))
+	print df.shape
+	# Specify all household variable names that is related
+	df = df[df['PUMA'].isin(map(int,pumaInArea))][['SERIALNO','PUMA','NP','VEH','HHT',
+	'HINCP','HUPAC','WIF']]
+	print df.shape
+	df.to_csv('data/ss15htx_clean.csv', index = False)
+
+	return df
+
+
+def getMarginal(dirname, filename, relevantSubjects):
+
+	'''
+	relevantSubjects specify names of variables we need
+	marginalDF is MultiIndex pandas DataFrame
+
+	'''
+
+	marginal = dict()
+	rawTable = pd.read_csv(os.path.join(dirname,filename), skiprows = [0])
+
+	# delete all columns which is just Margin of Error, and delete Female&Male subcategory
+	colNameToDrop = [cname for cname in rawTable.columns if cname.find('Margin of Error')>-1]
+	rawTable.drop(colNameToDrop,1,inplace = True)
+	colNameToDrop = [cname for cname in rawTable.columns[3:] if not cname.split(';')[0]=='Total']
+	rawTable.drop(colNameToDrop,1,inplace = True)
+
+	nrow, ncol = rawTable.shape
+
+	for i in range(nrow):
+		subjects = []
+		rownames = []
+		numbers = []
+		county = rawTable['Geography'][i]
+		for j in range(3,ncol):
[…]

### 2017-10-17 pop-synth/ipf.py

@@ -1,0 +1,109 @@
+from ipfn import *
+import numpy as np
+import pandas as pd
+from categorize import *
+import sys
+from datetime import datetime
+
+
+
+def getIPFresult_for_county(marginal1d, marginal2d, joint_dist, 
+	tolerance = 1e-3, max_iterations = 500, jd_zero_sub = 0.001):
+
+	'''
+	the "marginals" is marginal distributions of varaibles in the level of county
+	joint_dist should be joint distribution generated from sample data in a county 
+
+	marginals - MultiIndex pandas Series
+	joint_dist - MultiIndex pandas Series developed from product
+
+	'''
+	# joint distribution needs zero cells to be augmented
+	# zero cell problem
+	df = joint_dist.rename('total').reset_index()
+	df.replace(0, jd_zero_sub, inplace = True)
+
+	aggregates = []
+	dimensions = []
+	#for key in marginals.index.levels[0]:
+	#	aggregates.append(marginals[key])
+	for key in marginal1d.keys():
+		aggregates.append(marginal1d[key])
+		dimensions.append([key])
+	for key in marginal2d.keys():
+		aggregates.append(marginal2d[key])
+		dimensions.append([key[0], key[1]])
+
+	IPF = ipfn(df, aggregates, dimensions,convergence_rate = tolerance, max_iteration = max_iterations)
+	result = IPF.iteration()
+
+	return result
+	
+
+
+
+
+def setup_IPF_for_Ps(df, marginalDF, countyTable, mapCTtoPUMA, categories):
+
+	'''
+	Do IPF one by one in the level of county
+
+	'''
+
+	counties = marginalDF.columns
+
+	results = []
+	for county in counties:
+		countyID = countyTable[countyTable['name'] == county]['id']
+		pumaID_in_county = mapCTtoPUMA[mapCTtoPUMA['COUNTYFP'] == countyID]['PUMA5CE']
+		# select sample data and marginals for this county
+		sample_of_county = df[df['PUMA'].isin(pumaID_in_county)]
+		marginal_of_county = marginalDF[county]
+		# get joint distribution of sample data in the county
+
+		joint_dist_of_county = getPsJointDist(sample_of_county, categories)
+		county_result = getIPFresult_for_county(marginal_of_county, joint_dist_of_county)
+		result.append(county_result)
+
+	result = pd.concat(results, axis = 1)
+	return result
+
+
+
[…]

### 2017-10-17 pop-synth/marginal_processing.py

@@ -1,0 +1,255 @@
+import pandas as pd
+import os
+import itertools
+import numpy as np
+
+
+
+def getHhMarginalFrom2D(dirname, filename, subjects, one_marginal, two_marginal):
+	'''
+	subjects - two dimension list, the first subject should be the first hierarchy in the table
+	one_marginal - two level dict, the keys of first layer is county name, 
+					the keys for the second level is the attribute name,
+					values are 1D marginal distributions for those attributes in the county
+	two_marginal - two level dict, the first level is the same as 'one_marginal'
+					the second level uses 2D tuple as keys
+					values are 2d marginal distribution for two attributes in the key
+
+	'''
+
+	table = pd.read_csv(os.path.join(dirname, filename),skiprows = [0])
+
+	# Keep only columns with keywork "Estimate"
+	columnsToDrop = [col for col in table.columns[3:] if col.split(';')[0] != 'Estimate']
+	table.drop(columnsToDrop,1,inplace = True)
+	
+	nrow, ncol = table.shape
+
+	for i in range(nrow):
+		county = table['Geography'][i]
+		if county not in one_marginal.keys():
+			one_marginal[county] = {}
+			two_marginal[county] = {}
+		cat1, num1, cat2, num2, joint1, joint2, num = [],[],[],[],[],[],[]
+		for j in range(3,ncol):
+			colname = table.columns[j].split(';')[-1].strip()
+			# first level of 1D marginal distribution 
+			if colname.count(':')== 1 and colname.split(':')[-1] != '':
+				catname = colname.split(':')[-1]
+				cat1.append(catname[catname.find('-')+2:])
+				num1.append(table.iloc[i,j])
+			# second level of 1D marginal distribution 
+			elif colname.count(':') == 2 and colname.split(':')[-1] == '':
+				catname = colname.split(':')[-2]
+				cat2.append(catname[catname.find('-')+2:])
+				num2.append(table.iloc[i,j])
+			elif colname.count(':') == 2 and colname.split(':')[-1] != '':
+				c1, c2 = colname.split(':')[-2], colname.split(':')[-1]
+				if c2[c2.find('-')+2:] == '3 workers':
+					c2 = ' - 3 or more workers'
+				joint1.append(c1[c1.find('-')+2:])
+				joint2.append(c2[c2.find('-')+2:])
+				num.append(table.iloc[i,j])
+
+		
[…]

### 2017-10-17 pop-synth/pgm.R

@@ -1,0 +1,16 @@
+
+
+library(bnlearn)
+data("learning.test")
+bn.gs <- gs(learning.test)
+
+bn.hc <- hc(learning.test, score='aic')
+
+par(mfrow = c(1,2))
+plot(bn.gs, main='Grow-shrink algorithm')
+plot(bn.hc, main ='Hill Climbing')
+
+par(mfrow = c(1,2))
+highlight.opts <- list(nodes=c('A','B'), arcs = c('A','B'), col='red',fill='grey')
+graphviz.plot(bn.gs, main='gs',highlight = highlight.opts)
+graphviz.plot(bn.hc, main='hc', highlight = highlight.opts)

### 2017-10-17 pop-synth/viewTract.py

@@ -1,0 +1,9 @@
+from mpl_toolkits.basemap import Basemap
+import matplotlib.pyplot as plt
+
+
+map = Basemap(projection='cyl')
+map.fillcontinents(color = 'grey')
+
+plt.show()
+

### 2017-10-19 road-network-generator/extraFunctions.py
Generated SIMPLE road networks for Santiago, not including LARGE output files.

@@ -1,0 +1,164 @@
+import networkx as nx
+import numpy as np
+import csv,pdb,os,itertools,copy
+from collections import OrderedDict
+from numpy.linalg import norm
+from network import *
+
+
+def processAndAddConnections(nodes,tempnodeDict,links,segments,lanes,laneconnections,turninggroups,segToLink):
+    '''
+    '''
+    laneswithDwConn = []  # lanes with downstream connections
+    laneswithUpConn = []  # lanes with upstream connections
+
+    for connid,conn in laneconnections.iteritems():
+        fLane,tLane = conn.fromlane,conn.tolane
+        laneswithDwConn.append(fLane)
+        laneswithUpConn.append(tLane)
+    laneswithDwConn,laneswithUpConn = set(laneswithDwConn),set(laneswithUpConn)
+
+    totLaneList = lanes.keys()
+    terlaneswithoutDwConn, terlaneswithoutUpConn = set(),set()
+
+    #---Getting the terminal lanes ---lanes at the end and begining of links--- that are not connected to sink (source) node and no downstream (upstream) connection
+    for linkid,link in links.iteritems():
+        tnode,hnode = link.upnode,link.dnnode
+        #---first segment
+        segid = link.segments[0]
+        segment = segments[segid]
+        for i in xrange(segment.numlanes):
+            laneid = str(int(segid)*100 + i)
+            if laneid not in laneswithUpConn and tnode not in tempnodeDict['source']:
+                terlaneswithoutUpConn.add(laneid)
+        #---last segment
+        segid = link.segments[-1]
+        segment = segments[segid]
+        for i in xrange(segment.numlanes):
+            laneid = str(int(segid)*100 + i)
+            if laneid not in laneswithDwConn and hnode not in tempnodeDict['sink']:
+                terlaneswithoutDwConn.add(laneid)
+
+
+    #----Handling lanes without downstream connections
+    unresolvedLanes = []
+    connKeylist = laneconnections.keys()
+    count = len(connKeylist) + 1 # count and id of the connections
+    #---Copying the connections from one of the adjacent lanes
+    for laneid1 in terlaneswithoutDwConn:
+        resolved = False
+        segid1 = laneid1[:-2]
+        for i in xrange(segments[segid1].numlanes):
+            laneid2 = str(int(segid1)*100 + i)
+            if laneid1!=laneid2:
+                connkeys = [key for key in connKeylist if key[0]==laneid2]
+                if connkeys:
[…]

### 2017-10-19 road-network-generator/inputs/Arun-small-example/01AddSeqBase.py

@@ -1,0 +1,45 @@
+import sys, os
+import pdb
+#sys.path=os.getcwd()
+infilename='Base_poly.csv'
+outfilename='Base.csv'
+
+def editline(curline, lastline):
+   curline=curline.replace('"', "")
+   curline=curline.replace("\n","")
+   l1=curline.split(",")
+   l1.insert(1,l1[0])
+   l0=lastline.split(",")
+   curline = ",".join(l1)
+   #pdb.set_trace()
+   if l0[0] == l1[0]:
+      curline+=","+str(int(l0[4])+1)
+   else:
+      curline+=",1"
+      
+   return curline
+
+def main():
+
+   #open files
+   infile=open(infilename,"r")
+   outfile=open(outfilename, "w")
+   outfile.write("shapeid,ID,x,y,seq\n")
+
+   #first link
+   #infile.readline()
+   aline=infile.readline()
+   line0=aline
+   line0=line0.replace('"', "")
+   line0=line0.replace("\n",",1")
+   line0=line0.split(',')
+   line0.insert(1,line0[0])
+   line0 = ",".join(line0)
+   print line0
+   
+   for aline in infile.readlines():
+      newline=editline(aline,line0)
+      line0=newline
+      outfile.write(newline+"\n")
+
+main()

### 2017-10-19 road-network-generator/inputs/Arun-small-example/process/remove.py

@@ -1,0 +1,11 @@
+files = ["base_attr.csv","base_pl.csv","points.csv"]
+
+for file in files:
+    dummy = file.split(".")
+    file2 = dummy[0] + "2." + dummy[1]
+    with open(file,'r') as ifile:
+        with open(file2,'w') as ofile:
+            for aLine in ifile:
+                aLine = aLine.replace('"','')
+                ofile.write(aLine)
+        

### 2017-10-19 road-network-generator/inputs/handle-osmnet.py

@@ -1,0 +1,122 @@
+import pandas as pd
+import numpy as np
+import geopandas as gpd
+
+# To create
+# Base-attributes.csv
+# must have fields: shapeid, ID, NAME, NB_LANES, CAPACITY, SPEED, LENGTH, DIRECTION
+# FNODE, TNODE, LinkCat,
+
+# Base.csv
+# fields: shapeid,ID,x,y,seq
+
+# Points.csv
+# fields: shapeid,x,y,nodeid,trafficLid
+inputFolder = 'Santiago/from-OSMNET/'
+outputFolder = 'Santiago/mainConversion-Input/'
+
+def createBaseAttr():
+    necessary_cols = ['name','lanes','maxspeed','from','to','highway','oneway','distance' ]
+    edges = pd.read_csv(inputFolder + 'Santiago-Edges.csv', index_col=False, usecols=necessary_cols, low_memory=False)
+    edges.rename(columns={'name': 'NAME', 'lanes': 'NB_LANES', 'maxspeed': 'SPEED', 'from': 'FNODE',
+                'to': 'TNODE', 'highway': 'LinkCat', 'oneway': 'DIRECTION', 'distance': 'LENGTH'}, inplace=True)
+    edges.index.name = 'shapeid'
+    edges['ID'] = edges.index
+    edges['ctps_ID'] = edges.index # TODO
+
+    def set_direction(dir):
+        if dir == 'yes':
+            return -1 # two-way
+        else:
+            return 1
+    def set_nb_lanes(nb_lanes):
+        try:
+            return min(int(nb_lanes), 7)
+        except:
+            return 2
+    def set_LinkCat(LinkCat):
+        return 'A'
+
+    def set_SPEED(speed):
+        try:
+            return float(speed)
+        except:
+            return 60
+
+    edges['LinkCat'] = edges['LinkCat'].apply(set_LinkCat)
+    edges['SPEED'].fillna(60, inplace=True)
+    edges['NAME'].fillna('No-name', inplace=True)
+    edges['DIRECTION'] = edges['DIRECTION'].apply(set_direction)
+    edges['NB_LANES'] = edges['NB_LANES'].apply(set_nb_lanes)
+    edges['CAPACITY'] = edges['NB_LANES']*10
+    edges.to_csv(outputFolder + 'Base-attributes.csv')
+
+def createPoints():
[…]

### 2017-10-19 road-network-generator/mainConversion.py

@@ -1,0 +1,123 @@
+from processing import*
+from extraFunctions import*
+from collections import OrderedDict
+
+
+#inputFolder = "Inputs/Small3"
+#inputFolder = "Inputs/GBAFull"
+#inputFolder = "Inputs/GBAFullv3"
+# inputFolder = "Inputs/Small1"
+inputFolder = "Inputs/Santiago/mainConversion-Input/"
+
+#inputFolder = "Inputs/Kat"
+baseFname = os.path.join(inputFolder,"Base-shortened.csv")
+baseAttFname = os.path.join(inputFolder,"Base-attributes.csv")
+#baseAttFname = os.path.join(inputFolder,"subset.csv")
+pointFname = os.path.join(inputFolder,"Points.csv")
+connectFname = os.path.join(inputFolder,"Connections_Sumo.csv")
+typeToWidthFname = os.path.join(inputFolder,"LinkCat_Roadtype_LaneWidth.csv")
+ffsFname = os.path.join(inputFolder,"HCM2000.csv")
+
+#==================================================================================================
+
+def conv(s):
+    try:
+        s=float(s)
+    except ValueError:
+        pass
+    return s
+
+#==================================================================================================
+
+def readData(baseFname,baseAttFname,pointFname):
+    '''
+    '''
+    # reading base file
+    basePoly = OrderedDict()
+    with open(baseFname,'rb') as ifile:
+        reader = csv.DictReader(ifile)
+        for row in reader:
+            id = row['ID']
+            if id not in basePoly:
+                basePoly[id] = []
+            basePoly[id].append(row)
+
+    # sort basePoly
+
+    # reading base attributes
+    baseAttributes = OrderedDict()
+    with open(baseAttFname,'rb') as ifile:
+        reader = csv.DictReader(ifile)
+        for row in reader:
+            baseAttributes[row['ID']] = row
+
+    # reading points
+    points = OrderedDict()
+    with open(pointFname,'rb') as ifile:
+        reader = csv.DictReader(ifile)
+        for row in reader:
+            points[row['nodeid']] = row
+    #pdb.set_trace()
+
+    # reading connectionss
+    connSumo = OrderedDict()
+    # with open(connectFname,'rb') as ifile:
+    #     reader = csv.DictReader(ifile)
+    #     count = 0
+    #     for row in reader:
+    #         connSumo[(row['from_lane'],row['to_lane'])] = row
+    #         count += 1
+    #     if count>len(connSumo):
[…]

### 2017-10-19 road-network-generator/network.py

@@ -1,0 +1,309 @@
+import os,csv,pdb
+from operator import itemgetter
+
+class Network:
+    
+    def __init__(self,nodes,links,segments,lanes,laneconnections,turninggroups,linktts):
+        self.nodes = nodes
+        self.links = links
+        self.segments = segments # constructed at the final stage
+        self.lanes = lanes
+        self.laneconnections = laneconnections
+        self.turninggroups = turninggroups
+        self.linktts = linktts
+        self.numNodes = len(self.nodes)
+        self.numLinks = len(self.links)
+        self.numSegs = len(self.segments)
+    
+    def write(self,foldername):
+        #---Writing nodes
+        nodeFname = os.path.join(foldername,'node-nodes.csv')
+        with open(nodeFname,'wb') as ofile:
+            writer = csv.writer(ofile)
+            writer.writerow(['id','x','y','z','traffic_light_id','tags','node_type'])
+            nodeList = self.nodes.keys()
+            nodeList = sorted([int(item) for item in nodeList])
+            for id in nodeList:
+                node = self.nodes[str(id)]
+                writer.writerow(node.render())
+
+        #---Writing segments
+        segFname1 = os.path.join(foldername,'segment-attributes.csv')
+        segFname2 = os.path.join(foldername,'segment-nodes.csv')
+        with open(segFname1,'wb') as ofile1:
+            with open(segFname2,'wb') as ofile2:
+                writer1 = csv.writer(ofile1)
+                writer1.writerow(['id','link_id','sequence','num_lanes','capacity','max_speed','tags','link_category'])
+                writer2 = csv.writer(ofile2)
+                writer2.writerow(['id','x','y','z','seq_id'])
+                segList = self.segments.keys()
+                segList = sorted([int(item) for item in segList])
+                for id in segList:
+                    segment = self.segments[str(id)]
+                    writer1.writerow(segment.render())
+                    for point in segment.position:
+                        aList = [id,point['x'],point['y'],0,point['seq']]
+                        writer2.writerow(aList)
+
+        #---Writing segments
+        segFname1 = os.path.join(foldername,'segment-attributes_more.csv')
+        with open(segFname1,'wb') as ofile1:
+            writer1 = csv.writer(ofile1)
+            writer1.writerow(['id','link_id','sequence','num_lanes','capacity','max_speed','tags','link_category','length'])
+            segList = self.segments.keys()
+            segList = sorted([int(item) for item in segList])
[…]

### 2017-10-19 road-network-generator/processing.py

@@ -1,0 +1,613 @@
+import networkx as nx
+import numpy as np
+import csv,pdb,os,itertools,copy,math
+from collections import OrderedDict
+from numpy.linalg import norm
+from network import *
+from extraFunctions import *
+
+outputFolder = "Output"
+errorFolder = "Corrections"
+
+#=================================================================================================
+
+def getHCMSpeeds(filename):
+    '''
+        reads the HCM csv and returns a dictionary; check for duplicate keys
+    '''
+    ffsFunc = {}
+    with open(filename,'rb') as ifile:
+        reader = csv.DictReader(ifile)
+        for aRow in reader:
+            #print aRow
+            key = (aRow['LinkCat'],aRow['NumLane'],int(aRow['Capacity_vphPerLane']))
+            assert key not in ffsFunc
+            ffsFunc[key] = int(aRow['Speed_kmph'])
+    return ffsFunc
+
+
+
+#==================================================================================================
+
+
+def getFfs(LinkCat,numLanes,capacity,ffsFunc):
+    '''
+        returns the closed ffs from the HCM values in ffsFunc
+        noet that input capacity is total one
+    '''
+    if numLanes < 4:
+        adict = {k:v for k,v in ffsFunc.iteritems() if k[0]==LinkCat and k[1]==str(numLanes)}
+    else:
+        adict = {k:v for k,v in ffsFunc.iteritems() if k[0]==LinkCat and k[1]=='3+'}
+
+    capacities = [k[2] for k in adict]
+    closeCap = min(capacities, key=lambda x:abs(x-capacity/numLanes))
+
+    if numLanes < 4:
+        return ffsFunc[(LinkCat,str(numLanes),closeCap)]
+    else:
+        return ffsFunc[(LinkCat,'3+',closeCap)]
+
+
+
+
+#==================================================================================================
+
+def checkConnections(connAttributes,baseAttributes,tempnodeDict,lanes):
+    '''
+        Checks the connections so that they meet SimMobility's requirements
+    '''
+
+    # Getting the list of lanes that have no downstream connections or no upstream connections (and are not connected to source or sink nodes)
+    connList = connAttributes.keys()  # a list of all connections (fromlane,tolane)
+    laneswithDwConn = set([item[0] for item in connList])  # lanes with downstream connections
+    laneswithUpConn = set([item[1] for item in connList])  # lanes with upstream connections
+
+    connLanes = laneswithDwConn.union(laneswithUpConn)
+    laneSet = set([lane for lane in lanes])
[…]

### 2017-12-15 manuscript/manuscript.tex
revision to framework document

--- 2017-08-15 review/review.tex
@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
+\documentclass[11pt,twoside]{article}
+\usepackage{etex}
+\newcommand{\num}{6{} }
+
+\raggedbottom
+
+%geometry (sets margin) and other useful packages
+\usepackage{geometry}
+\geometry{top=1in, left=1in,right=1in,bottom=1in}
+ \usepackage{graphicx,booktabs,calc}
+
+%=== GRAPHICS PATH ===========
 \graphicspath{{./images/}}
-\usepackage{xcolor}
-\usepackage{amsmath}
+% Marginpar width
+%Marginpar width
+\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } } 
+\setlength{\marginparwidth}{.5in}
+%\reversemarginpar
+%\setlength{\marginparsep}{.02in}
+
+%% Fonts
+% \usepackage{fourier}
+% \usepackage[T1]{pbsi}
+
+\usepackage{lmodern}
+\usepackage[T1]{fontenc}
+%\usepackage{minted}
+
+\usepackage{rotating}
+%% Cite Title
+\usepackage[style=authoryear,maxcitenames=2,natbib]{biblatex}
+\addbibresource{../bib/references.bib}
+
+%%% Counters
+\usepackage{chngcntr,mathtools}
+\counterwithout{figure}{section}
+\counterwithout{table}{section}
+
+\numberwithin{equation}{section}
+
+%% Captions
+\usepackage{caption}
+\captionsetup{
+  labelsep=quad,
+  justification=raggedright,
+  labelfont=sc
+}
+
+%AMS-TeX packages
+\usepackage{amssymb,amsmath,amsthm} 
+\usepackage{bm}
+\usepackage[mathscr]{eucal}
+\usepackage{colortbl}
+\usepackage{color}
+
+\usepackage{epstopdf,subfigure,hyperref,enumerate,polynom,polynomial}
+\usepackage{multirow,minitoc,fancybox,array,multicol}
+
+\definecolor{slblue}{rgb}{0,.3,.62}%
+%\renewcommand{\chaptermark}[1]{ \markboth{#1}{} }
+\renewcommand{\sectionmark}[1]{ \markright{#1}{} }
+
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
+%\addtolength{\headwidth}{\marginparwidth}
+%\fancyheadoffset{30pt}
+%\fancyfootoffset{30pt}
+\fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
+\fancyhead[RO,LE]{\small Page \thepage} 
+\fancyfoot[RO,LE]{\small }% PR \num S-2015} 
+\fancyfoot[LO,RE]{\small }%\scshape MODL} 
+\cfoot{} 
+\renewcommand{\headrulewidth}{0.1pt} 
+\renewcommand{\footrulewidth}{0pt}
+%\setlength\voffset{-0.25in}
+%\setlength\textheight{648pt}
+
+\usepackage{paralist}
+
+%%% FORMAT PYTHON CODE
 \usepackage{listings}
-\usepackage[english]{babel}
-\usepackage{enumerate}
-\lstset{frame=tlrb,xleftmargin=\fboxsep,xrightmargin=-\fboxsep}
+% Default fixed font does not support bold face
+\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8} % for bold
+\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8}  % for normal
 
-\usepackage[style=numeric, backend=biber, maxcitenames=2]{biblatex}
-\bibliography{bib/references}
+% Custom colors
+\usepackage{color}
+\definecolor{deepblue}{rgb}{0,0,0.5}
+\definecolor{deepred}{rgb}{0.6,0,0}
+\definecolor{deepgreen}{rgb}{0,0.5,0}
 
+%\usepackage{listings}
+
+% % Python style for highlighting
+% \newcommand\pythonstyle{\lstset{
+% language=Python,
+% basicstyle=\footnotesize\ttm,
+% otherkeywords={self},             % Add keywords here
+% keywordstyle=\footnotesize\ttb\color{deepblue},
+% emph={MyClass,__init__},          % Custom highlighting
+% emphstyle=\footnotesize\ttb\color{deepred},    % Custom highlighting style
+% stringstyle=\color{deepgreen},
+% frame=tb,                         % Any extra options here
+% showstringspaces=false            % 
+% }}
+
+% % Python environment
+% \lstnewenvironment{python}[1][]
+% {
+% \pythonstyle
+% \lstset{#1}
+% }
+% {}
+
+% % Python for external files
+% \newcommand\pythonexternal[2][]{{
+% \pythonstyle
+% \lstinputlisting[#1]{#2}}}
+
+% % Python for inline
+% \newcommand\pythoninline[1]{{\pythonstyle\lstinline!#1!}}
+
+\newcommand{\osn}{\oldstylenums}
+\newcommand{\dg}{^{\circ}}
+\newcommand{\lt}{\left}
+\newcommand{\rt}{\right}
+\newcommand{\pt}{\phantom}
+\newcommand{\tf}{\therefore}
+\newcommand{\?}{\stackrel{?}{=}}
+\newcommand{\fr}{\frac}
+\newcommand{\dfr}{\dfrac}
+\newcommand{\ul}{\underline}
+\newcommand{\tn}{\tabularnewline}
+\newcommand{\nl}{\newline}
+\newcommand\relph[1]{\mathrel{\phantom{#1}}}
+\newcommand{\cm}{\checkmark}
+\newcommand{\ol}{\overline}
+\newcommand{\rd}{\color{red}}
+\newcommand{\bl}{\color{blue}}
+\newcommand{\pl}{\color{purple}}
+\newcommand{\og}{\color{orange!90!black}}
+\newcommand{\gr}{\color{green!40!black}}
+\newcommand{\nin}{\noindent}
+\newcommand{\la}{\lambda}
+\renewcommand{\th}{\theta}
+\newcommand{\al}{\alpha}
+\newcommand{\G}{\Gamma}
+\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
+            \node[shape=circle,draw,thick,inner sep=1pt] (char) {\small #1};}}
+
+\newcommand{\bc}{\begin{compactenum}[\quad--]}
+\newcommand{\ec}{\end{compactenum}}
+
+\newcommand{\p}{\partial}
+\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
+\newcommand{\dpd}[2]{\dfrac{\partial{#1}}{\partial{#2}}}
+\newcommand{\pdd}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
 
-\title{Population Synthesis Literature Review}
-\author{Jin Zhou}
+\title{Prototype city generation for mobility simulation in urban typologies: a  generalized framework}
+\author{}
 \maketitle
 
-\section{Goal}
-Since the step of urban typology has been done, we need to fill in the blank between city clustering results and a complete SimMobility\cite{adnan2016simmobility} input of virtual city.
-The city model that SimMobility takes as input has two components.
-One is the demand and the other is the supply, which includes the road and transit network.
-To model the demand, a methodology has to be developed to construct the population with desired demographics and allocate the population in space.
-Next to it, we also need to figure out how to use the synthesized households and individuals in disaggregate level to get the travel demand in the city.
-The literature review at this time is to gain a better idea of previous efforts to address the problem of population synthesis and assignment.
-
-\subsection{Questions:}
-\begin{enumerate}
-	\item What should the input of travel demand generation look like? This guides the adaptation of population synthesis method from literature.
-	\item What attributes should be included to characterize the population? Age, sex, educational level, and there must be some attributes reflect the traveling behavior of people. 
-	\item Do we synthesize households or personnel? Or how to match individuals to households?
-	\item We need to come up with a solution to cities don't have PUMS data and virtual cities that don't even have census data for demographics.	
-\end{enumerate}
-
-\section{Prior Work}
-
+\section{Introduction}
 Household assignment and spatial allocation are two big issues we need to solve when facing the problem of population synthesis.
@@
 
+\section{Background}
 There has been lots of literature on population synthesis.
@@
 
-The population synthesis method that \textcite{petrik2016measuring} used adapted ideas from probabilistic graphical models (PGM).
+The population synthesis method that %\cite{petrik2016measuring}
+used adapted ideas from probabilistic graphical models (PGM).
 In their graphical model, there are 7 socio-demographic attributes, which are Age, Gender, Employment, Income, Car ownership, Purpose of trip, and Who pay for the trip, to represent a person.
@@
 The flexibility of this method is that those conditional probabilities can be derived from different sources like census data or surveys on individuals.
+Nevertheless, compared to the high dependency of IPF on the quality of sample data, Bayesian method gives synthetic population more heterogeneity.
 Finally, the synthetic population is just random draws from the probabilistic graphical model.
@@
 \section{Data}
-Most paper in population synthesis used the Public Use Microdata Sample (PUMS), which is available both in individual level and household level.
+Most paper in population synthesis used Public Use Microdata Sample (PUMS) data, which is available both in individual level and household level.
 It includes multiple socio-economic parameters of a household or an individual in the region.
 The sample rate of this dataset is about 1\% to 10\% of the total population in the region. 
-We found PUMS data is complete in US cities but haven't found microdata in other cities in the world.
-Currently, we have found aggregated information of population, age, sex, household, education in districts in Beijing.
+PUMS data is very complete in US cities and can be easily downloaded from government census website \footnote{https://www.census.gov/programs-surveys/acs/data/pums.html}, but the same data for cities in developing countries can be relatively outdated and harder to find.
+The sample data is used as seeds to initialize the contingency table in IPF methods.
+Another important part of data is called aggregate data \footnote{https://factfinder.census.gov}.
+They give marginal distribution for variables of our interests.
+It's worth noticing that the marginal distributions for subsets with more than 2 variables can be rarely found.
+An ideal format for aggregate data is that we have the marginal distributions in the level of geographical zone cause IPF is carried out separately in each zone.
+So when downloading the aggregate data, one should first choose a geographic type in corresponding to the zones in the virtual city, then specify the variables or tables you want.
 
@@
 What is also important for us is the simplicity and novelty of methodology.
+We will start from a conventional IPF method using data from Houston city, then according to the execution time and the accuracy of the result we will make further extension to IPF method.
 
-\section{Iterative Proportional Fitting}
+\subsection{Network processing}
 
+\section{Method}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{l l l}
+    &&
+  \end{tabular}
+  \caption{List of symbols for variables and parameters}
+  \label{tab:symbols}
+\end{table}
+
+\begin{enumerate}[Step 1:]
+\item Network: modify OSM network file, make it legal input to SimMobility
+  \begin{enumerate}[1.]
+  \item links has to be unidirectional
+  \item shorten the links and add turning points
+  \item maintain the attributes of each object in the network
+  \end{enumerate}
+\item Population Synthesis
+  \begin{enumerate}[1.]
+  \item Data collection: sample data, aggregate data for zones in each city
+  \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
+  \item Apply IPF algorithm and use the weights result to draw synthetic population
+  \item Assign households in each county to locations based on land use information
+  \end{enumerate} 
+\end{enumerate}
+
+\subsection{Supply}
+
+\subsection{Demand}
+
+Although applying IPF to population synthesis is prevailing and the idea is straightforward, the algorithm itself is time consuming and expansive once the combinations of demographics gets large.
+Suppose we have 2 gender categories, 7 age groups, 5 income groups, 5 levels of education, 9 types of occupation, 2 kinds of car ownership, and 5 household types.
+These alone give rise to $31\,500$ types of individuals, not to mention describing individuals with more demographics.
+Another issue with published methods is that the time we spend to generate population for one zone has to be multiplied by the number of zones in a city in order to get the whole population for the city, which means the time for a complete process is hundreds of that for a single zone if the zone is in the sense of census tract.
+Since what we want is something less time consuming, a new methodology has to be developed.
+
+Considering the accessibility of the two sources of data that the synthesizer needs, one can always get sampled survey data in a city level but not necessarily census data for zones in a city.
+Given the global scope of our problem, there are cities whose zones are even not defined properly.
+Thus, we will face the problem of defining the zones of a city by our own in case of the data is in shortage.
+In developing the zones in a city, we will borrow the idea of zones from \textcite{fielbaum2016parametric}, where a city is divided into one CBD and subcenters along with their peripheral.
+It has to be admitted that in this way, we will not have aggregate census data for zones. This is a problem we will address later.
+
+\subsubsection{Zonal disaggregation and random sampling}
+\textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
+First, they obtain land-use data, which they use to disaggregate zonal characteristics (e.g.\ household location or firm location) into raster cells.
+The size of these cells are chosen based on computational tractability and desired resolution.
+Weights are then assigned to the cells based on the corresponding land-use category.
+For the household example, \textcite{moekel2003microsimulation} aggregate land-use attributes into the following: high-density housing, low-density housing, industry and open space;
+to which weights of 10, 5, 1 and 1 are assigned, respectively.
+The weights are then normalized by zone, and households are randomly selected to populate each cell to fit each proportion.
+The address of each household is assumed to be the centroid of each cell, which each associated with its parent zone.
+The rasterization is performed by a so-called ``point-in-polygon'' algorithm.
+This approach presents a fast way for household assignment, assuming prior zonal assignment is available.
+The populations themselves can first be generated using a simple IPF/U routine.
+
+\subsection{Summary of framework}
+In our first exploration, we have used 5 attributes to describe a households and they are household type, household size, household income in the last 12 months, number of vehicles and number of workers in the household.
+We not only find one dimensional marginal distribution for those variables but also two dimensional marginal distributions for household type $\times$ household size, number of workers $\times$ household size as well as number of vehicles $\times$ number of workers in the household.
+Across the process of data retrieving and processing, we find the categories each variable can have is largely depends on the availability and format of aggregate data.
+
+\section{Results and discussion}
+
+\section{Conclusion}
+\section{References}
+\printbibliography
+
+\appendix
+\section{Further notes on hierarchical iterative proportional fitting}
 Using IPF to synthesize population includes two major steps.
@@
 In the next step, the synthetic population shall be constructed by selecting households from PUMS in proportion to the estimated weights in the multiway table.
+Or we can do Monte Carlo sampling method to draw households from estimated weights from IPF.
 Specifically, the number of households of each demographic type in a census tract can be obtained by multiplying the total number of households by the probability in the cell representing this household type, or by drawing the numbers at random according to these probabilities.
@@
 
-\section{Proposed method}
-
-\begin{enumerate}[Step 1:]
-\item Network: modify OSM network file, make it legal input to SimMobility
-  \begin{enumerate}[1.]
-  \item links has to be unidirectional
-  \item shorten the links and add turning points
-  \item maintain the attributes of each object in the network
-  \end{enumerate}
-\item Population Synthesis
-  \begin{enumerate}[1. ]
-  \item Data collection: survey data, aggregate census data for zones in each city
-  \item Learn the parameter and the structure of Bayesian Network which captures the structure of population
-  \item sampling from Bayesian Network to get whole population in the city
-  \item Clustering the synthetic population in a city by K-means
-  \item Divide the city into zones according to clustering result, and form the aggregate census data for zones
-  \end{enumerate} 
-\end{enumerate}
-
-Although applying IPF to population synthesis is prevailing and the idea is straightforward, the algorithm itself is time consuming and expansive once the combinations of demographics gets large.
-Suppose we have 2 gender categories, 7 age groups, 5 income groups, 5 levels of education, 9 types of occupation, 2 kinds of car ownership, and 5 household types.
-These alone give rise to $31\,500$ types of individuals, not to mention describing individuals with more demographics.
-Another issue with published methods is that the time we spend to generate population for one zone has to be multiplied by the number of zones in a city in order to get the whole population for the city, which means the time for a complete process is hundreds of that for a single zone if the zone is in the sense of census tract.
-Since what we want is something less time consuming, a new methodology has to be developed.
-
-Considering the accessibility of the two sources of data that the synthesizer needs, one can always get sampled survey data in a city level but not necessarily census data for zones in a city.
-Given the global scope of our problem, there are cities whose zones are even not defined properly.
-Thus, we will face the problem of defining the zones of a city by our own in case of the data is in shortage.
-In developing the zones in a city, we will borrow the idea of zones from \textcite{fielbaum2016parametric}, where a city is divided into one CBD and subcenters along with their peripheral.
-It has to be admitted that in this way, we will not have aggregate census data for zones. This is a problem we will address later.
-
-\subsection{Two approaches to spatial assignment of households}
-\subsubsection{K-means clustering}
-With regard to the assignment of populations to zones, we will cluster the population at city level with K-means first.
-We reason that similar and highly correlated people will most likely live in the same area within a city.
-Then we obtain grouped people with similar attributes.
-We choose K-means because it is more intuitive if we want a spatial allocation and the feature of hidden centroid for each cluster will hopefully capture the characteristic of the zone.
-
-\subsubsection{Zonal disaggregation and random sampling}
-\textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
-First, they obtain land-use data, which they use to disaggregate zonal characteristics (e.g.\ household location or firm location) into raster cells.
-The size of these cells are chosen based on computational tractability and desired resolution.
-Weights are then assigned to the cells based on the corresponding land-use category.
-For the household example, \textcite{moekel2003microsimulation} aggregate land-use attributes into the following: high-density housing, low-density housing, industry and open space;
-to which weights of 10, 5, 1 and 1 are assigned, respectively.
-The weights are then normalized by zone, and households are randomly selected to populate each cell to fit each proportion.
-The address of each household is assumed to be the centroid of each cell, which each associated with its parent zone.
-The rasterization is performed by a so-called ``point-in-polygon'' algorithm.
-This approach presents a fast way for household assignment, assuming prior zonal assignment is available.
-The populations themselves can first be generated using a simple IPF/U routine.
-
-\newpage
-\section{References}
-\printbibliography[heading=none]
 \end{document}

### 2017-12-15 review/review.tex
added review documents to tree

@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
+\documentclass[11pt,twoside]{article}
+\usepackage{etex}
+\newcommand{\num}{6{} }
+
+\raggedbottom
+
+%geometry (sets margin) and other useful packages
+\usepackage{geometry}
+\geometry{top=1in, left=1in,right=1in,bottom=1in}
+ \usepackage{graphicx,booktabs,calc}
+
+%=== GRAPHICS PATH ===========
 \graphicspath{{./images/}}
-\usepackage{xcolor}
-\usepackage{amsmath}
+% Marginpar width
+%Marginpar width
+\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } } 
+\setlength{\marginparwidth}{.5in}
+%\reversemarginpar
+%\setlength{\marginparsep}{.02in}
+
+%% Fonts
+% \usepackage{fourier}
+% \usepackage[T1]{pbsi}
+
+\usepackage{lmodern}
+\usepackage[T1]{fontenc}
+%\usepackage{minted}
+
+\usepackage{rotating}
+%% Cite Title
+\usepackage[style=authoryear,maxcitenames=2,natbib]{biblatex}
+\addbibresource{bib/references.bib}
+
+%%% Counters
+\usepackage{chngcntr,mathtools}
+\counterwithout{figure}{section}
+\counterwithout{table}{section}
+
+\numberwithin{equation}{section}
+
+%% Captions
+\usepackage{caption}
+\captionsetup{
+  labelsep=quad,
+  justification=raggedright,
+  labelfont=sc
+}
+
+%AMS-TeX packages
+\usepackage{amssymb,amsmath,amsthm} 
+\usepackage{bm}
+\usepackage[mathscr]{eucal}
+\usepackage{colortbl}
+\usepackage{color}
+
+\usepackage{epstopdf,subfigure,hyperref,enumerate,polynom,polynomial}
+\usepackage{multirow,minitoc,fancybox,array,multicol}
+
+\definecolor{slblue}{rgb}{0,.3,.62}%
+%\renewcommand{\chaptermark}[1]{ \markboth{#1}{} }
+\renewcommand{\sectionmark}[1]{ \markright{#1}{} }
+
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
+%\addtolength{\headwidth}{\marginparwidth}
+%\fancyheadoffset{30pt}
+%\fancyfootoffset{30pt}
+\fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
+\fancyhead[RO,LE]{\small Page \thepage} 
+\fancyfoot[RO,LE]{\small }% PR \num S-2015} 
+\fancyfoot[LO,RE]{\small }%\scshape MODL} 
+\cfoot{} 
+\renewcommand{\headrulewidth}{0.1pt} 
+\renewcommand{\footrulewidth}{0pt}
+%\setlength\voffset{-0.25in}
+%\setlength\textheight{648pt}
+
+\usepackage{paralist}
+
+%%% FORMAT PYTHON CODE
 \usepackage{listings}
-\usepackage[english]{babel}
-\usepackage{enumerate}
-\lstset{frame=tlrb,xleftmargin=\fboxsep,xrightmargin=-\fboxsep}
+% Default fixed font does not support bold face
+\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8} % for bold
+\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8}  % for normal
 
-\usepackage[style=numeric, backend=biber, maxcitenames=2]{biblatex}
-\bibliography{bib/references}
+% Custom colors
+\usepackage{color}
+\definecolor{deepblue}{rgb}{0,0,0.5}
+\definecolor{deepred}{rgb}{0.6,0,0}
+\definecolor{deepgreen}{rgb}{0,0.5,0}
 
+%\usepackage{listings}
+
+% % Python style for highlighting
+% \newcommand\pythonstyle{\lstset{
+% language=Python,
+% basicstyle=\footnotesize\ttm,
+% otherkeywords={self},             % Add keywords here
+% keywordstyle=\footnotesize\ttb\color{deepblue},
+% emph={MyClass,__init__},          % Custom highlighting
+% emphstyle=\footnotesize\ttb\color{deepred},    % Custom highlighting style
+% stringstyle=\color{deepgreen},
+% frame=tb,                         % Any extra options here
+% showstringspaces=false            % 
+% }}
+
+% % Python environment
+% \lstnewenvironment{python}[1][]
+% {
+% \pythonstyle
+% \lstset{#1}
+% }
+% {}
+
+% % Python for external files
+% \newcommand\pythonexternal[2][]{{
+% \pythonstyle
+% \lstinputlisting[#1]{#2}}}
+
+% % Python for inline
+% \newcommand\pythoninline[1]{{\pythonstyle\lstinline!#1!}}
+
+\newcommand{\osn}{\oldstylenums}
+\newcommand{\dg}{^{\circ}}
+\newcommand{\lt}{\left}
+\newcommand{\rt}{\right}
+\newcommand{\pt}{\phantom}
+\newcommand{\tf}{\therefore}
+\newcommand{\?}{\stackrel{?}{=}}
+\newcommand{\fr}{\frac}
+\newcommand{\dfr}{\dfrac}
+\newcommand{\ul}{\underline}
+\newcommand{\tn}{\tabularnewline}
+\newcommand{\nl}{\newline}
+\newcommand\relph[1]{\mathrel{\phantom{#1}}}
+\newcommand{\cm}{\checkmark}
+\newcommand{\ol}{\overline}
+\newcommand{\rd}{\color{red}}
+\newcommand{\bl}{\color{blue}}
+\newcommand{\pl}{\color{purple}}
+\newcommand{\og}{\color{orange!90!black}}
+\newcommand{\gr}{\color{green!40!black}}
+\newcommand{\nin}{\noindent}
+\newcommand{\la}{\lambda}
+\renewcommand{\th}{\theta}
+\newcommand{\al}{\alpha}
+\newcommand{\G}{\Gamma}
+\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
+            \node[shape=circle,draw,thick,inner sep=1pt] (char) {\small #1};}}
+
+\newcommand{\bc}{\begin{compactenum}[\quad--]}
+\newcommand{\ec}{\end{compactenum}}
+
+\newcommand{\p}{\partial}
+\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
+\newcommand{\dpd}[2]{\dfrac{\partial{#1}}{\partial{#2}}}
+\newcommand{\pdd}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
@@
 What is also important for us is the simplicity and novelty of methodology.
+We will start from a conventional IPF method using data from Houston city, then according to the execution time and the accuracy of the result we will make further extension to IPF method.
+
+\section{Network processing}
+\subsection{OSM download}
+Currently we use MapZen to download the metro area network.
+We still have to isolate only relevant polygons however.
+
+\subsection{GIS2FE}
+We use this tool to convert \texttt{...\_osm\_line.shp} files to a network.
+This tool was developed by \cite{karduni2016protocol,}.
 
@@
 \item Population Synthesis
-  \begin{enumerate}[1. ]
-  \item Data collection: survey data, aggregate census data for zones in each city
-  \item Learn the parameter and the structure of Bayesian Network which captures the structure of population
-  \item sampling from Bayesian Network to get whole population in the city
-  \item Clustering the synthetic population in a city by K-means
-  \item Divide the city into zones according to clustering result, and form the aggregate census data for zones
+  \begin{enumerate}[1.]
+  \item Data collection: sample data, aggregate data for zones in each city
+  \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
+  \item Apply IPF algorithm and use the weights result to draw synthetic population
+  \item Assign households in each county to locations based on land use information
   \end{enumerate} 
@@
 It has to be admitted that in this way, we will not have aggregate census data for zones. This is a problem we will address later.
+
+\subsection{Data processing and IPF}
+In detail, the data processing and IPF includes following steps: Getting marginal distribution of variables that we are interested in from Census Bureau; Processing the data of marginal distributions to convenient format; Getting sample data from PUMS, and selecting sample data in our metro area by mapping census tract to county; Selecting variables in sample data, categorizing sample data based on "PUMS codebook" and variables categories in aggregate data; Generating joint distribution of sample data, which means get the frequency of each kind of combination of characteristics in sample data; Using joint distribution of sample data and marginal distribution as input to IPF; Performing IPF algorithm; Using output joint distribution of IPF to draw synthetic population.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.5\textwidth]{ipf_flow_chart.png}
+	\caption{Flow chart of data processing and IPF}
+	\label{fig: flowchart}
+\end{figure}
+
+In our first exploration, we have used 5 attributes to describe a households and they are household type, household size, household income in the last 12 months, number of vehicles and number of workers in the household.
+We not only find one dimensional marginal distribution for those variables but also two dimensional marginal distributions for household type $\times$ household size, number of workers $\times$ household size as well as number of vehicles $\times$ number of workers in the household.
+Across the process of data retrieving and processing, we find the categories each variable can have is largely depends on the availability and format of aggregate data.
 
@@
 \newpage
-\section{References}
-\printbibliography[heading=none]
+%\section{References}
+\printbibliography
 \end{document}

### 2017-12-21 manuscript/manuscript.tex
typology table

@@
 
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{l l l}\toprule
+    \bf SN &\bf Name &\bf Supply candidate \\\midrule
+    1 & Auto Sprawl & Baltimore \\
+    2 & Innovative Heavyweights & Madrid \\
+    3 & Sustainable Anchors & Vancouver \\
+    4 & Congested Middleweights & Bandung \\
+    5 & Boomers & Lagos \\
+    6 & Emerging Green Anchors & Hiroshima \\
+    7 & Dense Clique & Isfahan\\
+    8 & Hybrid Congested & Mashdad \\
+    9 & BRT-{\it STAR}s & Sao Paulo  \\
+    10 & Metro-Bike Giants & Dalian \\ \bottomrule
+  \end{tabular}
+  \caption{Prototype city descriptions and candidates}
+  \label{tab:proto}
+\end{table}
+
 \begin{enumerate}[Step 1:]

### 2017-12-21 manuscript/manuscript.tex
od cost estimation method

@@
 
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{l l l l l }
+    City/Typology & Microdata & Aggregate data & Land use & Transit \\
+  \end{tabular}
+  \caption{Data availability for candidate cities}
+  \label{tab:avail}
+\end{table}
+
 \begin{enumerate}[Step 1:]
@@
 The populations themselves can first be generated using a simple IPF/U routine.
+
+\subsection{Estimation of origin-destination matrices}
+There are three key
+
+\begin{table}[h!]\small
+  \centering
+  \begin{tabular}{l l l } \toprule
+    \textbf{Parameters}            & \textbf{Description}  & \textbf{Notes} \\ \midrule
+    \texttt{distance}     & Zone-zone distance (km) & \\
+    \texttt{car\_cost\_erp} & Road pricing cost (monetary units) &  \\
+    \texttt{car\_ivt}     & Car in-vehicle time  (hours) & \\
+    \texttt{pub\_ivt}     & Public transit in-vehicle time (hours) & \\
+    \texttt{pub\_walkt}   & Public transit access-egress walking time (hours) & \\
+    \texttt{pub\_wtt}     & Public transit waiting time (hours) & \\
+    \texttt{pub\_cost}    & Public transit cost between zones (monetary units) & \\
+    \texttt{avg\_transfer}& Average number of public transit transfers between zones & \\
+    \texttt{pub\_out}     & Public transit out-of-vehicle travel time (hours) & Sum  \texttt{pub\_walkt} and  \texttt{pub\_wtt} \\ \bottomrule
+  \end{tabular}
+  \caption{Operating cost parameters (Skim matrix elements) in SimMobility}
+  \label{tab:opcosts}
+\end{table}
+
+\subsubsection{Initial estimates}
+\begin{enumerate}
+\item The zone-zone distance can be readily estimated by averaging the shortest-path distances for all unique OD pairings within a pair of zones.
+Given $P_{n_O n_D}$ as the shortest path for a given node-node OD pair, $n_On_D$, between two zones $O$ and $D$, then we can write
+  \begin{equation}
+    d_{OD} = \frac{\sum_{n_On_D \in V(O) \times V(D)}|| P_{n_On_D}||}{|V(O) \times V(D)|}
+  \end{equation}
+\item The road pricing cost can be zero for now, but these will be relevant for congestion pricing scenarios. Thus,
+  \begin{equation}
+    C^{rp}_{OD} = 0
+  \end{equation}
+\item Car in-vehicle time $CIVT$ can be computed using free-flow speeds $FFS_e$ along each edge $e$ of the shortest-path ODs for each zone pair to travel times, which are then averaged to obtain the final estimate:
+  \begin{align}
+    CIVT(P) &= \sum_{e \in P} \fr{||e||}{FFS_e}\\
+    CIVT_{OD} &= \frac{\sum_{n_On_D \in V(O) \times V(D)} CIVT(P_{n_On_D})}{|V(O) \times V(D)|}
+  \end{align}
+\item Public transit in-vehicle time $PIVT$ is estimated using the procedure outlined in (iii) applied to the public transit graph
+\item  Public transit access-egress 
+\end{enumerate}
 

### 2017-12-21 manuscript/manuscript.tex
finalize od estimation procedure

@@
 \subsection{Estimation of origin-destination matrices}
-There are three key
+There are three key inputs at the zonal level required for running SimMobility preday: zonal attribute data, skim matrices, and zone-zone travel time data.
 
@@
 \item Public transit in-vehicle time $PIVT$ is estimated using the procedure outlined in (iii) applied to the public transit graph
-\item  Public transit access-egress 
+\item Public transit access-egress times are obtained by averaging the shortest edges connecting each road node to a public transit node in zone $O$ and adding this to the corresponding value for zone $D$.
+We assume a walking speed of 5 km/h, and dividing the sum obtained by this value produces an estimate of the walking time for each zone $OD$ pair.
+
+\item To estimate the public transit waiting time, we consider the median headway of the train/bus route on the first edge of all the shortest paths connecting public transit nodes in each zone-zone pairing.
+If there is a transfer on a given path, then the headway of the train/bus on the edge incident on the transfer point is added to the headway for the beginning of that path.
+
+\item The public transit cost between zone $OD$s is taken as the average fare scaled by the relative average length of the shortest paths between the nodes in a given pair of zones.
+
+\item We can compute the average number of transfers for each zone $OD$ by summing the number of transfers for the shortest paths in all node-node pairs between zones and dividing by the number of node-node pairings.
 \end{enumerate}

### 2017-12-26 manuscript/manuscript.tex
subsections on land use added

@@
 
-\begin{enumerate}[Step 1:]
-\item Network: modify OSM network file, make it legal input to SimMobility
-  \begin{enumerate}[1.]
-  \item links has to be unidirectional
-  \item shorten the links and add turning points
-  \item maintain the attributes of each object in the network
-  \end{enumerate}
-\item Population Synthesis
-  \begin{enumerate}[1.]
-  \item Data collection: sample data, aggregate data for zones in each city
-  \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
-  \item Apply IPF algorithm and use the weights result to draw synthetic population
-  \item Assign households in each county to locations based on land use information
-  \end{enumerate} 
-\end{enumerate}
+% \begin{enumerate}[Step 1:]
+% \item Network: modify OSM network file, make it legal input to SimMobility
+%   \begin{enumerate}[1.]
+%   \item links has to be unidirectional
+%   \item shorten the links and add turning points
+%   \item maintain the attributes of each object in the network
+%   \end{enumerate}
+% \item Population Synthesis
+%   \begin{enumerate}[1.]
+%   \item Data collection: sample data, aggregate data for zones in each city
+%   \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
+%   \item Apply IPF algorithm and use the weights result to draw synthetic population
+%   \item Assign households in each county to locations based on land use information
+%   \end{enumerate} 
+% \end{enumerate}
+
+In summary, the following are steps involving in generating a synthetic population and preparing it for mesoscopic simulation.
+\begin{itemize}
+\item 
+\item 
+\end{itemize}
 
@@
 
-\subsection{Estimation of origin-destination matrices}
+\subsection{Geolocation and spatial assignment}
+
+\subsubsection{Households}
+\begin{itemize}
+  \item Aggregate land use categories into these: low residential ($L$), high residential ($H$), commercial ($C$), industrial ($I$), education ($E$), open land ($O$)
+  \item Assign weights for households as follows:
+    \begin{equation}
+     HH (w_L, w_H, w_C, w_I, w_E, w_O) = (8, 10, 4, 1, 0, 0)
+    \end{equation}
+  \item Grid the map and assign weights to the cell $w_c^{HH}$ given its  prevailing land use category
+  \item Normalize cell weights $p_{c(SAL)}^{HH}$ in each second administrative level ($SAL$)
+    \begin{equation}
+      p_{c(SAL)}^{HH} = \fr{w_c^{HH}}{\sum_{c \in SAL} w_c^{HH}}
+    \end{equation}
+  \item Number of households in each cell given by:
+    \begin{equation}
+     N_{HH}^{c(SAL)} = p_c^{SAL}\cdot N_{HH}^{SAL}
+    \end{equation}
+    where $ N_{HH}^{SAL}$ is the number of households in each $SAL$.
+  \item Randomly sample to locate households in cell centroids for all $SAL$
+  \end{itemize}
+
+\subsubsection{Firms and education zones}
+\begin{itemize}
+  \item We obtain numbers of firms and schools in each $SAL$
+  \item Assign weights as follows:
+    \begin{align}
+      WORK (w_L, w_H, w_C, w_I, w_E, w_O) &= (1, 2, 10, 5, 3, 1)\\
+      EDU (w_L, w_H, w_C, w_I, w_E, w_O) &= (0, 0, 0, 0, 1, 0)
+    \end{align}
+  \item Assign weights to cells for work and education: $w_c^{WORK}$,  $w_c^{EDU}$
+
+  \item Find $p_{c(SAL)}^{WORK}$ and $p_{c(SAL)}^{EDU}$ as before
+  \item Find  $N_{WORK}^{c(SAL)}$ and $ N_{EDU}^{c(SAL)}$ as before
+  \item The final outputs here are the numbers of workplaces/schools in each TAZ:
+    \begin{align}
+      N_{WORK}^{TAZ(SAL)} &= \sum_{c(SAL) \in TAZ(SAL)} N_{WORK}^{c(SAL)} \\
+      N_{EDU}^{TAZ(SAL)} &= \sum_{c(SAL) \in TAZ(SAL)} N_{EDU}^{c(SAL)}  
+    \end{align}
+  \end{itemize}
+
+\subsection{Zonal-level  estimation}
+\subsubsection{Estimation of origin-destination matrices}
 There are three key inputs at the zonal level required for running SimMobility preday: zonal attribute data, skim matrices, and zone-zone travel time data.

### 2017-12-26 manuscript/manuscript.tex
brief entries on assignment and congested travel times

@@
 Suppose we have 2 gender categories, 7 age groups, 5 income groups, 5 levels of education, 9 types of occupation, 2 kinds of car ownership, and 5 household types.
-These alone give rise to $31\,500$ types of individuals, not to mention describing individuals with more demographics.
-Another issue with published methods is that the time we spend to generate population for one zone has to be multiplied by the number of zones in a city in order to get the whole population for the city, which means the time for a complete process is hundreds of that for a single zone if the zone is in the sense of census tract.
-Since what we want is something less time consuming, a new methodology has to be developed.
-
-Considering the accessibility of the two sources of data that the synthesizer needs, one can always get sampled survey data in a city level but not necessarily census data for zones in a city.
-Given the global scope of our problem, there are cities whose zones are even not defined properly.
-Thus, we will face the problem of defining the zones of a city by our own in case of the data is in shortage.
-In developing the zones in a city, we will borrow the idea of zones from \textcite{fielbaum2016parametric}, where a city is divided into one CBD and subcenters along with their peripheral.
-It has to be admitted that in this way, we will not have aggregate census data for zones. This is a problem we will address later.
 
@@
 
+\subsection{Trip generation and population assigment}
+In this step, we estimate the number of home-based work trips and home-based education trips between zones in order to have a starting point from which to assign individuals to usual work and education zones.
+The result is an OD matrix of trips from TAZ to TAZ.
+From this, we then randomly assign each individual to a zone for work or education.
+The method is here is an entropy maximization model (analogous to the gravity model).
+
+We also solve a static traffic assignment problem in order to find congested zone-zone travel times.
+The input network is a graph in which the nodes are TAZes and the edges are generalized between each zone.
+An average link capacity is obtained from the limiting capacity of the union of shortest paths between the zones.
+
+The congested travel time estimates for each TAZ OD is given by
+\begin{equation}
+  \label{eq:tt}
+  tt^{OD} = tt_0^{OD}\lt(1 + \lt( \fr{f_{OD}}{Cap_{OD}}\rt)^2\rt)
+\end{equation}
+where $tt_0^{OD}$ is the free-flow travel time for a given OD.
+
 \subsection{Zonal-level  estimation}
@@
 \end{enumerate}
+
+We assume morning peak period as 7-9am and evening peak period as 5-7 pm.
 

### 2017-12-26 manuscript/manuscript.tex
entry on entropy max model formulation
revised geolocation subsection

@@
 
-Although applying IPF to population synthesis is prevailing and the idea is straightforward, the algorithm itself is time consuming and expansive once the combinations of demographics gets large.
-Suppose we have 2 gender categories, 7 age groups, 5 income groups, 5 levels of education, 9 types of occupation, 2 kinds of car ownership, and 5 household types.
-
-\subsubsection{Zonal disaggregation and random sampling}
+\subsection{Geolocation and spatial assignment}
 \textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
@@
 This approach presents a fast way for household assignment, assuming prior zonal assignment is available.
-The populations themselves can first be generated using a simple IPF/U routine.
-
-\subsection{Geolocation and spatial assignment}
+Our implementation is outlined in the following subsections.
 
@@
 The method is here is an entropy maximization model (analogous to the gravity model).
+
+\begin{align}
+  \max & - \sum_{i=1}^K\sum_{j=1}^L x_{ij} \ln x_{ij}\\
+  \text{s.t.} & \sum_{j=1}^{L}x_{ij} = O_i, \quad i = 1,2, \dots,K\\
+       & \sum_{i=1}^{K}x_{ij} = D_j, \quad i = 1,2, \dots,L\\
+       &  \sum_{i=1}^K\sum_{j=1} c_{ij}x_{ij} = C \\
+      % &  \sum_{i=1}^K\sum_{j=1} c_{ij}x_{ij} = C \\
+       & x_{ij} \ge 0, \quad i =1,2,\dots,K,\; j  =1,2,\dots,L
+\end{align}
+
+The reader is referred to \citet{fang1995linearly} for a more formal background of this problem.
+A fuzzy extension that accounts for uncertainty in cost estimates has also been proposed by \citet{li2011entropy}.
+For now, however, we use the simple gravity approach outlined above.
+A quadratic cost constraint may also be introduced later:
+\begin{equation}
+   \sum_{i=1}^K\sum_{j=1} d_{ij}x_{ij}^2 = D \\
+\end{equation}
 

### 2017-12-26 manuscript/manuscript.tex
route assignment section added

@@
 From this, we then randomly assign each individual to a zone for work or education.
-The method is here is an entropy maximization model (analogous to the gravity model).
+
+\subsubsection{Trip distribution}
+The method for trip distribution is an entropy maximization model (analogous to the gravity model when there are no quadratic costs).
 
@@
 
+\subsubsection{Route assignment}
 We also solve a static traffic assignment problem in order to find congested zone-zone travel times.
@@
 An average link capacity is obtained from the limiting capacity of the union of shortest paths between the zones.
+The problem is a user-optimum (Wardorp) equilibrium, formulated as follows:
 
+\begin{align}
+  \min & \sum_{e_{OD} \in E} \int_{0}^{f_{OD}}tt^{OD}(x) dx \\ \notag
+  \text{s.t.} & \\
+  f_{OD} &= \sum_{i}\sum_{j}\sum_{r}\alpha_{ij}^{e_{OD}r}x_{ij}^{r}\\
+  \sum_r x_{ij}^{r} &= x_{ij} \\
+  f_{OD} &\ge 0 \\
+  x_{ij}^{r}& \ge 0 \\
+  \alpha_{ij}^{e_{OD}r} &=
+                          \begin{cases}
+                            1 & e_{OD} \in r \\
+                            0 & \text{otherwise}
+                          \end{cases}
+\end{align}
+
+The solution to this problem provides the equilbrium OD flows between the TAZes.
+We can therefore compute congestion travel times.
 The congested travel time estimates for each TAZ OD is given by
@@
   \label{eq:tt}
-  tt^{OD} = tt_0^{OD}\lt(1 + \lt( \fr{f_{OD}}{Cap_{OD}}\rt)^2\rt)
+  tt^{OD}(f_{OD}) = tt_0^{OD}\lt(1 + 0.15\lt( \fr{f_{OD}}{Cap_{OD}}\rt)^4\rt)
 \end{equation}
 where $tt_0^{OD}$ is the free-flow travel time for a given OD. 
+This is the standard Bureau of Public Roads congestion function \citep{mtoi2014calibration}, which we consider sufficient for the purposes of estimating congested travel times.
 

### 2017-12-27 manuscript/manuscript.tex
shell for framework diagram
tier diagram completed

@@
 \usepackage{paralist}
+
+%%%TIKZ
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+\usepackage{pgfgantt}
+\pgfplotsset{compat=newest}
+
+\usetikzlibrary{arrows,shapes,positioning,shapes.geometric}
+\usetikzlibrary{decorations.markings}
+\usetikzlibrary{shadows,automata}
+\usetikzlibrary{patterns}
+\usetikzlibrary{trees,mindmap,backgrounds}
+%\usetikzlibrary{circuits.ee.IEC}
+\usetikzlibrary{decorations.text}
+% For Sagnac Picture
+\usetikzlibrary{%
+    decorations.pathreplacing,%
+    decorations.pathmorphing%
+}
+\tikzset{no shadows/.style={general shadow/.style=}}
+%
+%\usepackage{paralist}
 
@@
 
-The population synthesis method that %\cite{petrik2016measuring}
-used adapted ideas from probabilistic graphical models (PGM).
+The population synthesis method that \citet{petrik2016measuring} used adapted ideas from probabilistic graphical models (PGM).
 In their graphical model, there are 7 socio-demographic attributes, which are Age, Gender, Employment, Income, Car ownership, Purpose of trip, and Who pay for the trip, to represent a person.
@@
 \section{Method}
-
 \begin{table}[h!]
@@
 
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}[>=latex]
+    \tikzset{tier/.style= {draw, rectangle, rounded corners, text width=25ex, align=center} }
+    \tikzset{source/.style= {fill=gray!50!, rectangle,text width=15ex,font={\small}} }
+    \tikzset{loca/.style= {draw,dotted, rectangle,text width=18ex,font={\small}} }
+    \tikzset{head/.style= {draw=none,text width=18ex,font={\footnotesize\it}},align=center }
+
+    \node[tier] (cell) at (0,0) {{\bf CELL}\\{\footnotesize via Rasterization}} ;
+    \node[tier] (taz) [above= of cell] {{\bf TAZ}\\{\footnotesize Traffic Analysis Zone} } ;
+    \node[tier] (sal) [above= of taz] {{\bf SAL} \\{\footnotesize Second Administrative Level}} ;
+    \node[tier] (utp) [above= of sal] {{\bf UTP} \\{\footnotesize Urban Typology Prototype}} ;
+
+    \node[head] (tier) [above= .5cm of utp] {Tier};
+    \node[head] (source) [right= 1cm of tier] {Source};
+    \node[head] (location) [right= 1cm of source] {Agent Location Resolution};
+
+    \path[draw, ->] (utp) -- (sal);
+    \path[draw, ->] (sal) -- (taz);
+    \path[draw, ->] (taz) -- (cell);
+
+    \node[source] (utp-source) [right= of utp] {Data};
+    \node[source] (sal-source) [right= of sal] {Data};
+    \node[source] (taz-source) [right= of taz] {Data/Generated};
+    \node[source] (cell-source)[right= of cell]{Generated};
+
+    \node[] (utp-location) [right= of utp-source] {};
+    \node[] (sal-location) [right= of sal-source] {};
+    \node[loca] (taz-location) [right= of taz-source] {Work (WORK), Education (EDU)};
+    \node[loca] (cell-location)[right= of cell-source]{Household (HH)};
+  \end{tikzpicture}
+  \caption{The four tiers of the prototype city development framework. The UTP represents the urban typology, and is given by the metro area of the candidate city chosen for the supply.
+  The Second Administrative Levels are the highest-level regions into which the UTP is subdivided.
+TAZ's are obtained for the UTP given the candidate road network.
+Where unavailable, these are generated as Thiessen polygons.
+CELLs are obtained by gridding over the SALs (or TAZs).
+The corresponding land use weight at the CELL level is used to determine the  {\it number} and {\it location} of the households therein; as well as the {\it numbers} of WORK and EDU points in each TAZ.}
+  \label{fig:framework}
+\end{figure}
 \subsection{Geolocation and spatial assignment}
@@
 
+\subsection{Calibration and validation}
+Initially tune parameters.
+
+\subsubsection{Trip structure}
+Aggregate modeshares for the typology
+
+\subsubsection{Activity structure}
+Consider cities near supply-demand interaction centroid:
+\begin{itemize}
+\item Supply-demand interaction defined by factors, namely: congestion/highway domination and inefficiency
+\item If travel surveys are available: validate against activity shares
+\end{itemize}
+
 \subsection{Summary of framework}

### 2017-12-27 manuscript/manuscript.tex
reorganized document

@@
 \section{Introduction}
-Household assignment and spatial allocation are two big issues we need to solve when facing the problem of population synthesis.
-Specifically, the first one addresses the problem of how to match individuals to the household where they belong, while the spatial allocation distributes households in geographical space.
-It is necessary because the ultimate goal of our work is to do transportation simulation, which can only be achieved by knowing the location of each agent. \\
+We describe in detail an automated pipeline for generating urban typology prototypes.
+These prototypes are simulation-ready environments that represent outcomes for the urban typologies discovered from global data collected in the first phase of the Mobility of the Future project.
+The organization of this document is as follows: First, a brief background and motivation is provided in Section 2. Subsequently, we describe our data sources and organization.
+In Section 4, the overall framework and submethods are described, with relationships clearly defined.
+Results are then presented and discussed in Section 5.
+Finally, we conclude with a summary and layout steps for further work.
+
+The contribution of this work lies in its highly generalized approach and its reliance only on open data.
+There are several examples in the extant literation on the synthesis and generation of cities for agent-based mobility simulation.
+However, these are mostly city-specific.
+
+Here, we combine existing methods in creating a pipeline that produces prototypes that represent the heterogeneity of urban structure and outcomes.
+The outputs from our approach serve as inputs in our state-of-the-art integrated urban laboratory---SimMobility \citep{adnan2016simmobility}, which we will use to evaluate future scenarios across these prototypes.
 
 \section{Background}
+{\it Too population-synthesis-heavy!! Needs REVISION; some material might need to go to the appendix}\\
+
 There has been lots of literature on population synthesis.
@@
 
+\textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
+First, they obtain land-use data, which they use to disaggregate zonal characteristics (e.g.\ household location or firm location) into raster cells.
+The size of these cells are chosen based on computational tractability and desired resolution.
+Weights are then assigned to the cells based on the corresponding land-use category.
+For the household example, \textcite{moekel2003microsimulation} aggregate land-use attributes into the following: high-density housing, low-density housing, industry and open space;
+to which weights of 10, 5, 1 and 1 are assigned, respectively.
+The weights are then normalized by zone, and households are randomly selected to populate each cell to fit each proportion.
+The address of each household is assumed to be the centroid of each cell, which each associated with its parent zone.
+The rasterization is performed by a so-called ``point-in-polygon'' algorithm.
+This approach presents a fast way for household assignment, assuming prior zonal assignment is available.
+
 \section{Data}
+{\it NEEDS REVISION!}\\
 Most paper in population synthesis used Public Use Microdata Sample (PUMS) data, which is available both in individual level and household level.
@@
 
-\subsection{Network processing}
-
-\section{Method}
+\subsection{Urban typologies}
 \begin{table}[h!]
   \centering
-  \begin{tabular}{l l l}
-    &&
-  \end{tabular}
-  \caption{List of symbols for variables and parameters}
-  \label{tab:symbols}
-\end{table}
-
-\begin{table}[h!]
-  \centering
+  \caption{Prototype city descriptions and candidates}
+  \label{tab:proto}
   \begin{tabular}{l l l}\toprule
@@
   \end{tabular}
-  \caption{Prototype city descriptions and candidates}
-  \label{tab:proto}
 \end{table}
@@
   \centering
+  \caption{Data availability for candidate cities}
+  \label{tab:avail}
   \begin{tabular}{l l l l l }
@@
   \end{tabular}
-  \caption{Data availability for candidate cities}
-  \label{tab:avail}
 \end{table}
 
-% \begin{enumerate}[Step 1:]
-% \item Network: modify OSM network file, make it legal input to SimMobility
-%   \begin{enumerate}[1.]
-%   \item links has to be unidirectional
-%   \item shorten the links and add turning points
-%   \item maintain the attributes of each object in the network
-%   \end{enumerate}
-% \item Population Synthesis
-%   \begin{enumerate}[1.]
-%   \item Data collection: sample data, aggregate data for zones in each city
-%   \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
-%   \item Apply IPF algorithm and use the weights result to draw synthetic population
-%   \item Assign households in each county to locations based on land use information
-%   \end{enumerate} 
-% \end{enumerate}
-
-In summary, the following are steps involving in generating a synthetic population and preparing it for mesoscopic simulation.
-\begin{itemize}
-\item 
-\item 
-\end{itemize}
-
-\subsection{Supply}
-
-\subsection{Demand}
+\section{Method}
 
@@
 \end{figure}
+
+\begin{table}[h!]
+  \centering
+  \begin{tabular}{l l l}
+    &&
+  \end{tabular}
+  \caption{List of symbols for variables and parameters}
+  \label{tab:symbols}
+\end{table}
+
+% \begin{enumerate}[Step 1:]
+% \item Network: modify OSM network file, make it legal input to SimMobility
+%   \begin{enumerate}[1.]
+%   \item links has to be unidirectional
+%   \item shorten the links and add turning points
+%   \item maintain the attributes of each object in the network
+%   \end{enumerate}
+% \item Population Synthesis
+%   \begin{enumerate}[1.]
+%   \item Data collection: sample data, aggregate data for zones in each city
+%   \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
+%   \item Apply IPF algorithm and use the weights result to draw synthetic population
+%   \item Assign households in each county to locations based on land use information
+%   \end{enumerate} 
+% \end{enumerate}
+
+In summary, the following are steps involving in generating a synthetic population and preparing it for mesoscopic simulation.
+\begin{itemize}
+\item 
+\item 
+\end{itemize}
+
+\subsection{Supply}
+The supply pipeline can be summarized in the following steps:
+\begin{itemize}
+\item Obtain candidate city for urban typology based on that closest to centroid of supply factors
+\item Process road network from supply candidate and generate SimMobility inputs
+\item Generate public transit graph from GTFS and other sources
+\item Process land-use data polygons
+\end{itemize}
+
+\subsection{Demand}
+
 \subsection{Geolocation and spatial assignment}
-\textcite{moekel2003microsimulation} present an intuitive approach to assigning households to zones, which we propose to implement.
-First, they obtain land-use data, which they use to disaggregate zonal characteristics (e.g.\ household location or firm location) into raster cells.
-The size of these cells are chosen based on computational tractability and desired resolution.
-Weights are then assigned to the cells based on the corresponding land-use category.
-For the household example, \textcite{moekel2003microsimulation} aggregate land-use attributes into the following: high-density housing, low-density housing, industry and open space;
-to which weights of 10, 5, 1 and 1 are assigned, respectively.
-The weights are then normalized by zone, and households are randomly selected to populate each cell to fit each proportion.
-The address of each household is assumed to be the centroid of each cell, which each associated with its parent zone.
-The rasterization is performed by a so-called ``point-in-polygon'' algorithm.
-This approach presents a fast way for household assignment, assuming prior zonal assignment is available.
 Our implementation is outlined in the following subsections.

### 2017-12-28 manuscript/manuscript.tex
completed demand-supply interactions

@@
 \definecolor{slblue}{rgb}{0,.3,.62}%
+
+\hypersetup{
+    colorlinks,%
+    citecolor=blue,%
+    filecolor=blue,%
+    linkcolor=blue,
+    urlcolor=slblue
+}
+
 %\renewcommand{\chaptermark}[1]{ \markboth{#1}{} }
@@
 \maketitle
+
+\tableofcontents
+
+\newpage
 
@@
 The corresponding land use weight at the CELL level is used to determine the  {\it number} and {\it location} of the households therein; as well as the {\it numbers} of WORK and EDU points in each TAZ.}
-  \label{fig:framework}
+  \label{fig:tiers}
 \end{figure}
@@
 
-In summary, the following are steps involving in generating a synthetic population and preparing it for mesoscopic simulation.
-\begin{itemize}
-\item 
-\item 
-\end{itemize}
-
 \subsection{Supply}
@@
 \subsection{Demand}
+
+\subsubsection{Synthesis}
+In summary, the following are steps involving in generating a synthetic population and preparing it for mesoscopic simulation.
+\begin{itemize}
+\item 
+\item 
+\end{itemize}
+
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}[>=latex]
+    \tikzset{data/.style= {fill=gray!50!, rectangle, text width=25ex, align=center} }
+    \tikzset{res/.style= {draw, rectangle, rounded corners,text width=18ex,font={\small}, align=center} }
+    \tikzset{exp/.style= {draw=none, rectangle,text width=18ex,font={\scriptsize}} }
+    %\tikzset{head/.style= {draw=none,text width=18ex,font={\footnotesize\it}},align=center }
+
+    \node[data] (aggr) at (0,0) {{\bf Aggregates (SAL)}\\{\footnotesize HH|IND|WORK|EDU}} ;
+    \node[data] (samp) [right= 3cm of aggr] {{\bf Samples (SAL)}\\{\footnotesize HH|IND} } ;
+    \node[res] (synhi) [below= 2cm of samp] {{\bf Synthetic Population} \\{\footnotesize HH|IND}} ;
+    \node[res] (synwe) [below= 2cm of aggr] {{\bf Synthetic Population} \\{\footnotesize WORK|EDU}} ;
+
+    \node[data, text width=15ex] (landu) [below right= of synwe] {LAND\\ USE};
+    \node[res] (locwe) [below left= of landu] {{\bf TAZ Locations}\\{\footnotesize WORK|EDU}};
+    \node[res] (lochi) [below right= of landu] {{\bf CELL Locations}\\{\footnotesize HH}};
+
+    \node[res] (assig) [below right= of locwe] {{\bf Assignment}\\{\footnotesize IND$\to$WORK|EDU}};
+
+    \path[draw, ->] (aggr) to [out=0,in=180] (synhi);
+    \path[draw, ->] (samp) -- node[exp, right]{Hierarchical Iterative Proportional Fitting} (synhi);
+    \path[draw, ->] (aggr) -- node[exp, left,align=right]{Iterative Proportional Fitting} (synwe);
+    \path[draw, ->] (synwe) to [out=270,in=180]   (landu);
+    \path[draw, ->] (synhi) to [out=270,in=0]   (landu);
+    \path[draw, ->] (landu) to [out=195,in=0]   (locwe);
+    \path[draw, ->] (landu) to [out=-15,in=180]   (lochi);
+    \path[draw, ->] (lochi) to [out=270,in=0]   (assig);
+    \path[draw, <-] (locwe) to [out=270,in=180] (assig);
+
+    \node[exp,align=center] (exp-assig) [above=.25cm of assig] {Entropy Maximization Model};
+    \node[exp,align=center] (exp-landu) [below=.25cm of landu] {rasterization \& weighting};
+    %\node[exp] (taz-source) [right= of synhi] {Hierarchical Iterative Proportional Fitting};
+    %\node[exp] (cell-source)[left= of synhi]{Generated};
+
+  \end{tikzpicture}
+  \caption{Demand-Supply Interaction piece of the prototype city generation framework}
+  \label{fig:framework}
+\end{figure}
 
@@
 \section{Conclusion}
-\section{References}
+%\section{References}
 \printbibliography

### 2017-12-28 manuscript/manuscript.tex
Demand diagram completed

@@
 
-\subsection{Zonal-level  estimation}
+\begin{figure}[h!]
+  \centering
+  \begin{tikzpicture}[>=latex]
+    \tikzset{pro/.style= {fill=gray!50!, rectangle, text width=30ex, align=center} }
+    \tikzset{res/.style= {draw, rectangle, rounded corners,text width=22ex,font={\small}, align=center} }
+    \tikzset{exp/.style= {draw=none, rectangle,text width=18ex,font={\scriptsize}} }
+    %\tikzset{head/.style= {draw=none,text width=18ex,font={\footnotesize\it}},align=center }
+
+    \node[res] (ff) at (0,0) {network freeflow parameters} ;
+    \node[pro] (trip) [below= of ff] {{\bf Trip Distribution}\\{\footnotesize static traffic assignment} } ;
+    \node[res] (tt) [below= of trip] {congested travel times} ;
+    \node[res] (est) [below= of tt] {peak \& off-peak network cost estimates} ;
+
+    % \node[pro] (pre) [below= 2cm of est] {run Pre-Day} ;
+    \node[pro] (calib) [below = of est] {{\bf Calibration \& Validation}\\{\footnotesize day-to-day learning}};
+    \node[res] (fin) [below = of calib] {final {\bf PreDay} and {\bf WithinDay} parameters};
+
+    \path[draw, ->] (ff) --(trip);
+    \path[draw, ->] (trip) -- (tt);
+    \path[draw, ->] (tt) -- (est);
+    \path[draw, ->] (est) to (calib);
+    %path[draw, ->] (pre) to [out=270,in=0]   (calib);
+    \path[draw, ->] (calib) to  (fin);
+    % \path[draw, ->] (lochi) to [out=270,in=0]   (assig);
+    % \path[draw, <-] (locwe) to [out=270,in=180] (assig);
+
+    % \node[exp,align=center] (exp-assig) [above=.25cm of assig] {Entropy Maximization Model};
+    % \node[exp,align=center] (exp-landu) [below=.25cm of landu] {rasterization \& weighting};
+    %\node[exp] (taz-source) [right= of synhi] {Hierarchical Iterative Proportional Fitting};
+    %\node[exp] (cell-source)[left= of synhi]{Generated};
+
+  \end{tikzpicture}
+  \caption{Demand piece of the prototype city generation framework}
+  \label{fig:framework}
+\end{figure}
+
 \subsubsection{Estimation of origin-destination matrices}

### 2018-01-09 manuscript/manuscript.tex
changed title
updates to manuscript

@@
 
-\title{Prototype city generation for mobility simulation in urban typologies: a  generalized framework}
-\author{}
+\title{A generalized framework for urban typology prototype generation for agent-based simulations in SimMobility}
+\author{Jimi Oke \and Iveel Tsogsuren \and Youssef Medhat \and Carlos Lima Azevedo \and Chris Zegras \and Joseph Ferreria \and Moshe Ben-Akiva}
+
 \maketitle
@@
 
-    \node[res] (assig) [below right= of locwe] {{\bf Assignment}\\{\footnotesize IND$\to$WORK|EDU}};
+    \node[res] (assig) [below =4cm of landu] {{\bf Assignment}\\{\footnotesize IND$\to$WORK|EDU}};
 

### 2018-01-11 synthetic-population/generate_pop.py
population synthesis and land use

@@ -1,0 +1,86 @@
+import pandas as pd
+import math
+import numpy as np
+
+counties = [24001, 24003, 24005, 24009, 24011, 24013, 24015, 24017, 24019, 24021, 24023, 24025, 24027, 24029, 24031, 24033, 24035, 24037, 24039, 24041, 24043, 24045, 24047, 24510]
+# counties = [24009, 24011, 24013, 24015, 24017, 24019, 24021, 24023, 24025, 24027, 24029, 24031, 24033, 24035, 24037, 24039, 24041, 24043, 24045, 24047, 24510]
+# [24001, 24003, 24005] #
+
+
+samples = pd.read_csv('samples.csv') # 52366, total weight: 187477.126042
+total_population = pd.read_csv('Baltimore_affs/ACS_15_5YR_B01003_with_ann.csv')
+
+def create_county_population(current_county, samples, total_population):
+    weights = pd.read_csv('Weights_multilevel_veh/' + current_county + '_multilevel_weights.csv')
+    population_row = total_population.loc[total_population['GEO.id2'] == current_county]
+    total = int(population_row.HD01_VD01.item())
+    print('_____________County: ', current_county, 'total population: ', total)
+    w_scaling = total / weights['weights'].sum()
+    samples['weight'] = weights['weights'] * w_scaling
+
+    # print('original weight total population: ', samples['weight'].sum()) # error 0
+
+    ######################### Integerize households #########################
+    unique_hh = samples.drop_duplicates('hhid', inplace=False)
+    counts = samples['hhid'].value_counts()
+    # assert (unique_hh['weight']*unique_hh['APER']).sum() - samples['weight'].sum() < 1
+
+    error_weight = 0
+    num_in_population = {}
+    syn_pop = 0
+
+    SMALL_HHID = {}
+    for index, row in unique_hh.iterrows():
+        if error_weight > 0:
+            int_num = math.floor(row.weight)
+        else:
+            int_num = math.ceil(row.weight)
+        error_weight += (int_num - row.weight) * row.APER # ind weight
+        syn_pop +=  int_num * row.APER
+        num_in_population[row.hhid] = int_num
+        SMALL_HHID[row.hhid] = index
+    print('synthetic total population', syn_pop) # (unique_hh['weight']*unique_hh['APER']).sum())
+
+    ####################### Write new population ############################
+    max_num = max(num_in_population.values())
+    scale = 10**(math.ceil(math.log(max_num, 10)))
+    hhsample = 0
+
+    new_population = []
[…]

### 2018-01-11 synthetic-population/generate_pop_all_attributes.py

@@ -1,0 +1,100 @@
+import pandas as pd
+import geopandas as gpd
+import math
+import numpy as np
+
+counties = [24001, 24003, 24005, 24009, 24011, 24013, 24015, 24017, 24019, 24021, 24023, 24025, 24027, 24029, 24031, 24033, 24035, 24037, 24039, 24041, 24043, 24045, 24047, 24510]
+# counties = [24009, 24011, 24013, 24015, 24017, 24019, 24021, 24023, 24025, 24027, 24029, 24031, 24033, 24035, 24037, 24039, 24041, 24043, 24045, 24047, 24510]
+# [24001, 24003, 24005] #
+
+Baltimore_Metro_Counties = [24003, 24510, 24005, 24013, 24027, 24035, 24025]
+
+samples = pd.read_csv('samples_with_all_attributes.csv') # 52366, total weight: 187477.126042
+total_population = pd.read_csv('Baltimore_affs/ACS_15_5YR_B01003_with_ann.csv')
+population_weights = '4-nonzero-point-weights/point_weights_'
+outFolder = 'Baltimore_syn_population_with_location/'
+
+def create_county_population(current_county, samples, total_population):
+    weights = pd.read_csv('Weights_multilevel_dec30/' + current_county + '_multilevel_weights.csv')
+    population_row = total_population.loc[total_population['GEO.id2'] == current_county]
+    total = int(population_row.HD01_VD01.item())
+    print('_____________County: ', current_county, 'total population: ', total)
+    w_scaling = total / weights['weights'].sum()
+    samples['weight'] = weights['weights'] * w_scaling
+
+    # print('original weight total population: ', samples['weight'].sum()) # error 0
+
+    ######################### Integerize households #########################
+    unique_hh = samples.drop_duplicates('hhid', inplace=False)
+    counts = samples['hhid'].value_counts()
+    # assert (unique_hh['weight']*unique_hh['APER']).sum() - samples['weight'].sum() < 1
+
+    error_weight = 0
+    num_in_population = {}
+    syn_pop = 0
+
+    SMALL_HHID = {}
+    for index, row in unique_hh.iterrows():
+        if error_weight > 0:
+            int_num = math.floor(row.weight)
+        else:
+            int_num = math.ceil(row.weight)
+        error_weight += (int_num - row.weight) * row.APER # ind weight
+        syn_pop +=  int_num * row.APER
+        num_in_population[row.hhid] = int_num
+        SMALL_HHID[row.hhid] = index
[…]

### 2018-01-11 synthetic-population/prepareHH.py

@@ -1,0 +1,136 @@
+import pandas as pd
+import csv
+import re
+folder = 'Baltimore_affs/ACS_15_5YR_'
+end = '_with_ann.csv'
+
+sex_age = folder + 'B01001' + end
+sex_education = folder + 'B15002' + end
+unweighted_sample_population_count = folder + 'B00001' + end
+unweighted_sample_housing = folder + 'B00001' + end
+total_population = folder + 'B01003' + end
+hhsize_vehicles = folder + 'B08201' + end
+hhtypes = folder + 'B09019' + end
+
+# Sex, age, education, size, vehicles
+
+# Outputs
+# group
+# metadata  gender: {HD01, HD02}
+outputFolder = 'Outputs/'
+gender_marg = outputFolder + 'GENDER.csv'
+age_marg = outputFolder + 'AGE.csv'
+edu_marg = outputFolder + 'EDU.dat'
+veh_marg = outputFolder + 'VEHICLES.dat'
+
+# sex_education = folder + 'ACS_15_5YR_B15002_SEX_BY_EDUCATIONAL.csv'
+# hhtype_size = folder + 'ACS_15_5YR_B11016_HOUSEHOLD TYPE BY HOUSEHOLD SIZE.csv'
+# hhincome = folder + 'ACS_15_5YR_B19001_HOUSEHOLD INCOME.csv'
+# df = pd.read_csv(sex_age_file, skiprows=1)
+# df = pd.read_csv(sex_education, skiprows=1)
+# df = pd.read_csv(hhtype_size, skiprows=1)
+# print("a_b".split("_")[1])
+# df = df.select(lambda x: not re.search('Margin of Error', x), axis=1)
+
+def cleanHeader(x):
+    attrs  = x.split(';')
+    if len(attrs) == 2:
+        return attrs[1]
+    return x
+
+def firstAttr(x):
+    attrs  = x.split(':')
+    if len(attrs) == 2:
+        return attrs[0]
+    return x
+
+def secondAttr(x):
+    attrs  = x.split(':')
+    if len(attrs) == 2:
+        return attrs[1]
+    return x
+
+def code_category(df, orderColumn):
+    attr_category = {}
+    attr_category_code = []
+    new_columns = []
+    for i in range(len(orderColumn)):
+        attr_category[str(i)] = orderColumn[i]
+        new_columns.append(str(i))
+    df.columns = new_columns # df.rename(columns = {orderColumn[i]:str(i)})
[…]

### 2018-01-11 synthetic-population/prepare_raw.py

@@ -1,0 +1,121 @@
+import pandas as pd
+
+# Prepare households
+# with open('usa_00001.csv','r') as f:
+#     for i in range(4):
+#         print(next(f))
+#     # head = [next(f) for x in range(5)]
+# # print(head)
+# marginal_files = ['ACS_15_5YR_B08202.csv','ACS_15_5YR_B08203.csv','ACS_15_5YR_B11016.csv','ACS_15_5YR_B19001.csv']
+
+
+# files names
+# sex_age = 'ACS_15_5YR_B01001.csv'
+# hsize_veh = 'ACS_15_5YR_B08201.csv'
+# workers_veh = 'ACS_15_5YR_B08203.csv'
+# htype_hsize = 'ACS_15_5YR_B11016.csv'
+# vw, vf, vn, wf, wn, fn = []
+
+# df = pd.read_csv('../data-input/marginals_2d.csv')
+
+# group_vehicles = {
+# 0:73165,
+# 1:93749,
+# 2:55933,
+# 3:12005,
+# 4:3548
+# }
+
+# "YEAR","DATANUM","SERIAL","HHWT","REGION","COUNTY","MET2013","CITYPOP","PUMA","CPUMA0010","GQ","OWNERSHP","OWNERSHPD","HHINCOME","UNITSSTR","PERNUM","PERWT","FAMSIZE","RELATE","RELATED","SEX","AGE","MARST","BIRTHYR","SCHOOL","EDUC","EDUCD","EMPSTAT","EMPSTATD","LABFORCE","OCC","INCTOT","FTOTINC","TRANWORK","TRANTIME"
+
+# COUNTY = 24510
+# FAMSIZE = people per household
+# sex =
+# age =
+# ownershp=?
+
+# df = pd.read_csv('usa_00002.csv')
+# # df.loc[df['COUNTY'] == 24510]]
+# # print(df.COUNTY)
+# # print(df[df.iloc[:,1].map(lambda x: type(x) == str)])
+# print(df.dtypes)
+# # df = df.where(df['COUNTY'] == 24510)
+# df = df[df.COUNTY == 24510]
+# # df = df[['FAMSIZE','SEX', 'AGE', 'OWNERSHP']]
+# df.to_csv("Balt.csv")
+
+# import csv
+# f = open('../data-input/hh_marginals_2d.csv')
+numLines = 0
+fromFile = open('usa_00003.csv', 'r')
+toFile = open('Baltimore_sample.csv', 'w')
+# header
[…]

### 2018-01-11 synthetic-population/statistics_syn.py

@@ -1,0 +1,25 @@
+import pandas as pd
+
+counties = [24001, 24003, 24005, 24009, 24011, 24013, 24015, 24017, 24019, 24021, 24023, 24025, 24027, 24029, 24031, 24033, 24035, 24037, 24039, 24041, 24043, 24045, 24047, 24510]
+# counties = [24001, 24003, 24005] # 24009, 24011, 24013, 24015, 24017, 24019, 24021, 24023, 24025, 24027, 24029, 24031, 24033, 24035, 24037, 24039, 24041, 24043, 24045, 24047, 24510]
+folder = 'Baltimore_syn_population2/'
+outputFolder = 'Stats/'
+vehicles_frames = []
+for county in counties:
+    current_county = str(county)
+    header = ['category', current_county]
+    vehicles = vehicles.drop_duplicates('hhid', inplace=True)
+    vehicles = pd.read_csv(folder + current_county + '_vehicles_stats.csv', header=None)
+
+
+    vehicles.columns = header
+    vehicles.set_index('category', inplace=True)
+    vehicles_frames.append(vehicles)
+
+
+vehicles = pd.concat(vehicles_frames, axis=1)
+vehicles.to_csv(outputFolder + 'agg_vehicles.csv', index=True)
+
+
+#################### Write totals ####################################
+vehicles.sum(axis=1).to_csv(outputFolder + 'Synthetic_sum_vehicles.csv', index=True)

### 2018-01-11 land-use/edu_location.py
number of schools in counties

@@ -1,0 +1,65 @@
+import pandas as pd
+import geopandas as gpd
+from shapely.geometry import Point
+
+inFolder = 'schools/'
+outFolder = 'school_locations/'
+
+high_edu = [
+'Higher_Education_Private_Four_Year',
+'Higher_Education_Private_Two_Year',
+'Higher_Education_Public_Four_Year',
+'Higher_Education_Public_Two_Year',
+]
+
+k_thru_12 = [
+'K_thru_12_Education_Charter_Schools',
+'K_thru_12_Education_Public_Schools',
+]
+
+counties = {
+    'AnneArundel': 'Anne_2010LULC', #24003
+    'BaltimoreCity': 'Baci_2010LULC', #24510
+    'BaltimoreCounty': 'Baco_2010LULC', #24005
+    'Carroll': 'Carr_2010LULC', #24013
+    'Howard' : 'Howa_2010LULC', #24027
+    'QueenAnne' : 'Quee_2010LULC', #24035
+    'Harford' : 'Harf_2010LULC' #24025
+}
+
+counties = ['Baltimore City','Baltimore', 'Anne Arundel', 'Howard', 'Carroll', "Queen Anne's", 'Harford']
+
+def get_k_thru_12():
+    charter = pd.read_csv(inFolder + 'K_thru_12_Education_Charter_Schools.csv', usecols=['X', 'Y', 'county'])
+    public = pd.read_csv(inFolder + 'K_thru_12_Education_Public_Schools.csv', usecols=['X', 'Y', 'County'])
+    public.rename(columns={"County": "county"}, inplace=True)
+    df = pd.concat([charter, public])
+    df = df.loc[df['county'].isin(counties)]
+    geometry = [Point(xy) for xy in zip(df.X, df.Y)]
+    df = df.drop(['X', 'Y'], axis=1)
+    crs = {'init': 'epsg:4326'}
+    gdf = gpd.GeoDataFrame(df, crs=crs, geometry=geometry)
+    gdf.to_file(outFolder+'k_thru_12_all')
+    print("k_thru_12")
+    print(gdf.groupby(['county']).size())
+
+def get_higher_edu():
+    # columns: ['X', 'Y', 'OBJECTID', 'sitename', 'sitedesc', 'streetaddr',
+    # 'building_s', 'city', 'county', 'state', 'zip', 'phonenumbe', 'tollfree_n',
[…]

### 2018-01-11 land-use/main.py

@@ -1,0 +1,62 @@
+import fiona
+import re
+import geopandas as gpd
+import baltimore_landuse_code
+
+counties = {
+    'AnneArundel': 'Anne_2010LULC', #24003
+    'BaltimoreCity': 'Baci_2010LULC', #24510
+    'BaltimoreCounty': 'Baco_2010LULC', #24005
+    'Carroll': 'Carr_2010LULC', #24013
+    'Howard' : 'Howa_2010LULC', #24027
+    'QueenAnne' : 'Quee_2010LULC', #24035
+    'Harford' : 'Harf_2010LULC' #24025
+}
+
+countyCodes = {
+    '24003': 'Anne_2010LULC', #24003
+    '24510': 'Baci_2010LULC', #24510
+    '24005': 'Baco_2010LULC', #24005
+    '24013': 'Carr_2010LULC', #24013
+    '24027' : 'Howa_2010LULC', #24027
+    '24035' : 'Quee_2010LULC', #24035
+    '24025' : 'Harf_2010LULC' #24025
+}
+
+def set_landuse_weights(inFolder='original_landuse_data/', outFile='weights/points/'):
+    for key, county in countyCodes.items():
+        landFile = inFolder + county + '/' + county + '.shp'
+        outFile = outFolder + key + '.shp'
+        df = gpd.read_file(landFile)
+        # print(df.head)
+        # print(df.LU_CODE.unique())
+        df['pop_weight'] = df.apply(lambda row: baltimore_landuse_code.HH_WEIGHTS[row.LU_CODE], axis=1)
+        df['edu_weight'] = df.apply(lambda row: baltimore_landuse_code.EDU_WEIGHTS[row.LU_CODE], axis=1)
+        df['firm_weight'] = df.apply(lambda row: baltimore_landuse_code.FIRM_WEIGHTS[row.LU_CODE], axis=1)
+        df['id'] = df.index + 1 # We need index starting from 1.
+        df.to_file(outFile)
+
+def clean_point_weights(inFolder='weights/points/', outFile='weights/clean/'):
+    for key, county in countyCodes.items():
+        raster_points = inFolder + key + '.shp'
+        outFile = outFolder + key + '.shp'
+        points = gpd.read_file(raster_points)
+        points = points[points.value != 0]
+        points.rename(columns={'value': 'TAZ'}, inplace=True)
+        points['TAZ'] = points['TAZ'].astype('int')
+        points.to_file(outFile)
+
[…]
